---
title: "Proliferation_ModID"
output: html_document
date: '2023-05-29'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Looking for modules in Proliferating CD4 

Much of the opioid literature suggests that proliferating CD4 are responsible for a large part of the opioid dependent CD4 impact. Objective here is to try and identify those modules. See module RMD for more of the intuition/logic as to choices made. 
```{r setup functions,echo=FALSE}
## Required packages for code
suppressPackageStartupMessages(require(WGCNA))
suppressPackageStartupMessages(require(flashClust))
suppressPackageStartupMessages(require(Hmisc))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(library(doParallel))
registerDoParallel(cores=10)
options(warn=0)

## Required packages for saving results
suppressPackageStartupMessages(require(openxlsx))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(cowplot))
suppressPackageStartupMessages(library(Seurat))
#these functions are from Sam Kazar's 2020 Nature Med paper with minor modifications and added notes
{# Choosing the appropriate power for generating the adjacency matrix.
  FindPower <- function(datExpr){
    #choose soft-threshold power
    powers = c(c(1:10), seq(from = 12, to=20, by=2))
    sft=pickSoftThreshold(datExpr,dataIsExpr = TRUE,powerVector = powers,corFnc = cor,
                          corOptions = list(use = 'p', method = "pearson"),networkType = "signed")
    
    # Plot the results
    par(mfrow = c(1,2));
    cex1 = 0.9;
    
    # Scale-free topology fit index as a function of the soft-thresholding power
    plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit, signed R^2",type="n", main = paste("Scale independence"));
    text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],labels=powers,cex=cex1,col="red");
    
    # Red line corresponds to using an R^2 cut-off
    abline(h=0.80,col="red")
    
    # Mean connectivity as a function of the soft-thresholding power
    plot(sft$fitIndices[,1], sft$fitIndices[,5],xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",main = paste("Mean connectivity"))
    text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
  }
  
  # Generating the adjacency matrix and performing clustering
  ClusterTOM <- function(datExpr, softPower){
    #dev.off()
    #Calclute the adjacency matrix
    adj= adjacency(datExpr,type = "signed", power = softPower);
    
    #Turn adjacency matrix into a topological overlap matrix (TOM) to minimize the effects of noise and spurious associations.
    TOM=TOMsimilarityFromExpr(datExpr,networkType = "signed", TOMType = "signed", power = softPower, corType="bicor");
    
    colnames(TOM) = rownames(TOM) = colnames(datExpr)
    dissTOM=1-TOM
    
    #Hierarchical clustering of the genes based on the TOM dissimilarity measure
    geneTree = flashClust(as.dist(dissTOM),method="complete");
    
    #Plot the resulting clustering tree (dendrogram)
    plot(geneTree, xlab="", sub="",cex=0.3);
    
    return(list(dissTOM = dissTOM, geneTree = geneTree)) #returns list with dissimilarity TOM, and the clustered gene tree.
  }
  
  # Cut the resulting clustering dendrogram using the "tree" method for cutreeDynamic. Minimum module size can be specified.
  CutTOMTree <- function(datExpr, dissTOM, geneTree, minModuleSize = 10){
    #dev.off()
    # Module identification using dynamic tree cut, you can also choose the hybrid method
    dynamicMods = cutreeDynamic(dendro = geneTree,  method="tree", minClusterSize = minModuleSize);
    #dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM, method="hybrid", deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = minModuleSize);
    
    #Get the module labels and the size of each module. Lable 0 is reserved for unassigned genes
    print(table(dynamicMods))
    
    #Plot the module assignment under the dendrogram; note: The grey color is reserved for unassigned genes
    dynamicColors = labels2colors(dynamicMods)
    table(dynamicColors)
    plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut", dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05, main = "Gene dendrogram and module colors")
    
    #Set the diagonal of the dissimilarity to NA 
    diag(dissTOM) = NA;
    
    #extract modules
    module_colors= setdiff(unique(dynamicColors), "grey")
    modules = lapply(module_colors, function(x){colnames(datExpr)[which(dynamicColors==x)]})
    names(modules) = module_colors
    return(list(dyanmicColors = dynamicColors, modules = modules)) #returns list with module colors, and the modules themselves
  }
  
  # Merge modules with low dissimilarity. Cutoff for dissimilarity merge can be specified
  MergeSimilarModules <- function(datExpr, dynamicColors, geneTree, MEDissThres = 0.5){
    #cacluate eigengenes
    MEList = moduleEigengenes(datExpr, colors=dynamicColors)
    MEs = MEList$eigengenes
    
    # Calculate dissimilarity of module eigengenes
    MEDiss = 1-cor(MEs);
    
    # Cluster module eigengenes
    METree = hclust(as.dist(MEDiss), method = "average");
    
    # Plot the result
    #sizeGrWindow(7, 6)
    plot(METree, main = "Clustering of module eigengenes",
         xlab = "", sub = "")
    abline(h = MEDissThres, lwd=2, col="red")
    
    # Call an automatic merging function
    merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)
    # The merged module colors
    mergedColors = merge$colors;
    # Eigengenes of the new merged modules:
    mergedMEs = merge$newMEs;
    #plot showing how merged modules exist
    plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                        c("Dynamic Tree Cut", "Merged dynamic"),
                        dendroLabels = FALSE, hang = 0.03,
                        addGuide = TRUE, guideHang = 0.05)
    
    #extract merged modules
    merged_module_colors= setdiff(unique(mergedColors), "grey")
    merged_modules = lapply(merged_module_colors, function(x){colnames(datExpr)[which(mergedColors==x)]})
    names(merged_modules) = merged_module_colors
    
    return(list(mergedColors = mergedColors, merged_modules = merged_modules)) #returns list with merged colors, and the merged modules themselves
  }
  
  ############################################################################################################
  ## Functions for assaying module significance against background and temporal variability in module score ##
  ############################################################################################################
  
  ## Test to determine if the genes within the module are truly the least dissimilar compared to randomly generated modules of the same size.
  TestModuleSignificance <- function(mod, dissTOM, expr.data, n_perm = 10000, pval = 0.05, n.bin = 10, seed = 42){
    #vectorize the actual distribution of (dis)similarities, and remove zeros!
    set.seed(seed)
    true.diss = as.vector(dissTOM[mod,mod])
    true.diss = true.diss[-which(true.diss == 0)]
    
    #size of module for permutations
    mod.size = length(mod)
    
    #bin all genes by expression
    expr.avg = rowMeans(expr.data)
    expr.avg = expr.avg[order(expr.avg)]
    expr.avg.cut = as.numeric(x = cut2(x = expr.avg, m=round(length(expr.avg)/n.bin)))
    names(expr.avg.cut) = names(expr.avg)
    
    #create a table of binnings of all genes and of our module genes
    all.bin.table = table(expr.avg.cut)
    mod.bin.table = table(expr.avg.cut[mod])
    
    #randomly generate module with same expression binning structure and run t.test and record results
    test.results = data.frame(statistic = rep(NA, n_perm), pval = rep(NA, n_perm)) #container for results
    
    for (i in 1:n_perm){ #by permutation
      random.mod = list() #create an empty list we will fill with gene names (each element will be gene names by bin)
      
      #for each bin of the mod bin table, randomly select that number of genes from the full set with that expression
      for (j in names(mod.bin.table)){ 
        bin.genes = sample(names(expr.avg.cut)[which(expr.avg.cut == as.numeric(j))], mod.bin.table[j], replace = FALSE)
        random.mod[[as.numeric(j)]] = bin.genes #stick those genes into the random.mod list
      }
      #unlist and vectorize the distribution of (dis)similarities (remember to remove zeros)
      random.mod = unlist(random.mod)
      random.diss = as.vector(dissTOM[random.mod,random.mod])
      random.diss = random.diss[-which(random.diss == 0)]
      
      #perform a one-sided wilcox.test and record the statistic and p-value.
      #Note, IMPORTANT: here we perform the test asking if the true diss is LESS THAN the random diss, as we are trying to minimize dissimilarity
      test = wilcox.test(x = true.diss, y = random.diss, alternative = "less")
      test.results[i,] = c(test$statistic, test$p.value)
    }
    
    #correct for multiple hypothesis testing, and then report the proportion of bad tests
    test.results$FDR = p.adjust(test.results$pval, method = "fdr")
    num.failed.tests = sum(test.results$FDR > pval)
    print(paste(paste(num.failed.tests, n_perm, sep="/"), "permutations failed the Mann-Whitney test.", sep=" "))
    
    #is the percentage of failed tests less than or equal to the p-val specified?
    return(num.failed.tests/n_perm <= pval) #returns a vector of booleans indicating if each module was significant based on the specific p.val
  }
  
 
}


#This function just wraps the above with appropriate prompts to identify modules 
#By default this is using the "RNA" assay. Modifying it is trivial but 
#default is set to protect against incorrectly using SCT 
ModID<-function(seuratobj, cluster=NA, prefix="modules", nPCS="", softpower=""){
  #subset to a cluster or clusters of interest if desired 
  if (sum(!is.na(cluster))>1){test_clus<-subset(seuratobj, idents=cluster)}else{test_clus<-seuratobj}
  #perform feature selection, scaling, and PCA 
  test_clus<-FindVariableFeatures(test_clus, verbose=FALSE)
  test_clus<-ScaleData(test_clus, verbose=FALSE)
  test_clus<-RunPCA(test_clus, verbose=FALSE)
  
  #prompot for number of PCs needed 
  if(nPCS==""){print(ElbowPlot(test_clus))}
  while(nPCS==""){nPCS<-readline("how many NPCs?")}
  nPCS<-as.integer(nPCS)
  #expanding the genes by projecting the dims to more PCs 
  test_clus<-ProjectDim(test_clus)
  
  #gere we'll grab the top 25 features (positive and negative) associated with each dimension 
  genes<-c()
  for (i in 1:nPCS){genelist<-TopFeatures(test_clus, dim = i, nfeatures = 50,balanced = T, projected = T )
  for (i in 1:length(genelist)){genes<-c(genes, genelist[[i]])}}
  #remove duplicates 
  genes<-unique(genes)
  #grab the matrix from the seurat object 
  test_clus<-as.matrix(test_clus@assays$RNA@data[genes,])
  #prompt for softpower 
  FindPower(datExpr=t(test_clus))
  while(softpower==""){softpower<-readline("what softpower?")}
  softpower<-as.integer(softpower)
  #run the remaining functions to actually cut the tree
  test_clus_tom<-ClusterTOM(datExpr = t(test_clus), softPower = softpower)
  test_clus_mods<-CutTOMTree(datExpr = t(test_clus), geneTree = test_clus_tom$geneTree, dissTOM = test_clus_tom$dissTOM, minModuleSize = 10 )
  test_clus_merge_mods<-MergeSimilarModules(datExpr=t(test_clus), dynamicColors = test_clus_mods$dyanmicColors, geneTree = test_clus_tom$geneTree, MEDissThres = 0.5)
  print(test_clus_merge_mods$merged_modules)                                          
  test_clus_merge_mods.isSig = sapply(test_clus_merge_mods$merged_modules, function(module){
    TestModuleSignificance(mod = module, dissTOM = test_clus_tom$dissTOM, expr.data = test_clus,
                           n_perm = 10000, pval = 0.05, n.bin = 10)
  })
  test_clus_merge_mods.isSig = test_clus_merge_mods$merged_modules[test_clus_merge_mods.isSig]
  print(test_clus_merge_mods.isSig)
  names(test_clus_merge_mods.isSig)<-paste(prefix,names(test_clus_merge_mods.isSig ), sep="_")
  return(test_clus_merge_mods.isSig)
}

#lets me wrap a whole group of objects
WrapGroupModID<-function(seuratobjlist, cluster, conds=""){mods<-list()
for (i in 1:length(seuratobjlist)){
  
  print(names(seuratobjlist)[[i]])
  if(length(conds)==1){res<-try(ModID(seuratobjlist[[i]],cluster=cluster, prefix=names(seuratobjlist)[[i]]))}else{
    res<-try(ModID(seuratobjlist[[i]],cluster=cluster, prefix=names(seuratobjlist)[[i]], nPCS =conds[i,1], softpower=conds[i,2]))
  }
  if(inherits(res, "try-error"))
  {
    #error handling code,  just skip this iteration if not modules get found or theres some error in the process
    next
  }
  
  
  mods[[i]]<-res}
return(mods)
}
#this function clusters mods and allows the user to select a cutoff for merging
MergeMods<-function(modslist,height=""){
  #first find the overlap bettween modules 
  modslist<-modslist[!unlist(lapply(modslist, is.null))]
  name<-c()
  for(i in 1:length(modslist)){
    name<-c(name, names(modslist[[i]]))
  }
  modslist<-unlist(modslist, recursive = FALSE)
  overlaps<-list()
  for(i in 1:length(modslist)){
    modoverlap<-list()
    for(j in 1:length(modslist)){
      modoverlap[[j]]<-length(intersect(modslist[[i]], modslist[[j]]))/length(modslist[[i]])

    }
    overlaps[[i]]<-modoverlap
  }
  
  #datawrangle that for plotting
  data<-matrix(unlist(overlaps), nrow = length(overlaps))
  rownames(data)<-colnames(data)<-name
  p1<-heatmap(data, keep.dendro=TRUE)
  print(heatmap(data))
  res2 = as.hclust( p1$Rowv )
  print(plot(res2))
  #ask for user cut 
  if(height==""){goodenough<-FALSE
  while(goodenough==FALSE){height<-readline(prompt ="where to make the cut")
  if(height==""){height=2}
  res<-cutree(res2, h=as.numeric(height))
  print("number of modules is")
  print(max(res))
  print(plot(res2))
  rect.hclust(res2, h=as.numeric(height), border=2:6)
  goodenough<-readline(prompt = "(w)rite mods or (c)ut again?")
  goodenough<-case_when(goodenough=="w"|goodenough=="W" ~TRUE, TRUE ~FALSE)
  }}else{
      res<-cutree(res2, h=as.numeric(height))
      rect.hclust(res2, h=as.numeric(height), border=2:6)
  }
  #merge based on those identified clusters
  mergedmods<-list()
  for( i in 1:max(res)){
    mergedmods[[i]]<-unique(unlist(modslist[res==i]))
  }
  #return modules 
  return(mergedmods)}



#loading data
results<-readRDS( "~/gibbs/DOGMAMORPH/Ranalysis/Objects/20230529CD4ObjAnno.rds")
#setting correct default
DefaultAssay(results)<-"RNA"

```

## Possible comparisons 

There are many iterations we could select for this analysis, from most ideal to least ideal: 
* Participant + Treatment + Timepoint 
* Treatment + Timepoint
* Participant + Treatment 
* Treatment 

Which can further be expanded to which CD4 populations are included. A more robust CD4 module ID will occur at a later point. First we just subset proliferating cells, then we get the parts of those that are in each subset. 

```{r, checking pops, fig.width=16, fig.height=9}

results<-subset(results, idents = "Proliferating")

table(results$Treatment)
table(results$Timepoint)

table(paste(results$Treatment, results$Timepoint))
table(paste(results$Treatment, results$Participant))
table(paste(results$Treatment, results$Timepoint, results$Participant))

```

We need "enough" cells to have a hope of identifying modules. "Enough" is a hard call, but in my experience, around 50 cells are needed to call a module. Clearly the most desirable (treatment, timepoint, and participant specific) is not sufficient to call modules. Treatment + pariticipant also has too few cells in the methadone and Naltrexone groups to ID modules. Treatment timepoint is borderline for all except methadone time 0, but seems close enough to try at least. Timepoint and treatment alone both have sufficient cells, and we can try to call modules with them. 

```{r, modid, fig.width=16, fig.height=9}

results$T_TP<-paste(results$Treatment, results$Timepoint, sep="_")

results.split<-SplitObject(results, "T_TP")
#npcs and softpowers, selecting npc based on last nonplataue pc on elbow plot, selecting softpower as first above 0.8 
#nothing for methadone, PCA doesnt have enough observations, go modules for the other two 
conds<-cbind(c(7,6, 7, 9,0,0),c(5,9, 8, 14,0,0 ))
mods1<-WrapGroupModID(results.split, "NA", conds)
saveRDS(mods1, "~/gibbs/DOGMAMORPH/Ranalysis/modules/20230529ProliferatingModules_Treatment_Timepoint.rds")

#just doing treatment, we get at least 2 per condition 
results.split<-SplitObject(results, "Treatment")
conds<-cbind(c(8, 7,6),c(6, 9,8))
mods2<-WrapGroupModID(results.split, "NA")
saveRDS(mods2, "~/gibbs/DOGMAMORPH/Ranalysis/modules/20230529ProliferatingModules_Treatment.rds")

#finally looking at all proliferating cells to see if we can find something interesting, just got 2 out of it
mods3<-ModID(results, nPCS = 9, softpower = 6)
saveRDS(mods3, "~/gibbs/DOGMAMORPH/Ranalysis/modules/20230529ProliferatingModules.rds")

#in these cases, it doesnt make a lot of sense to go through and merge modules based on similarity. We're trying to preserve differences, so we may want to be able to preserve these separations even if things end up harder to analyze in the short term 
```

```{r session info}
devtools::session_info()
```

