---
title: "20230525ModID"
output: html_document
date: '2023-05-25'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Module ID into 

Objective of this analysis is to identify networks of genes associated with disease states. Further work will annotate these. The below functions are modified from Kazer /et al./ 2020 /Nature Med/ and Collora /et al./ 2022 /Immunity/ 

note that these functions are being called with defined parameters to facilitate the rmd file knitting, but the functions are designed to work interactively without defining softpower or npcs. 
```{r setup functions}
## Required packages for code
suppressPackageStartupMessages(require(WGCNA))
suppressPackageStartupMessages(require(flashClust))
suppressPackageStartupMessages(require(Hmisc))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(library(doParallel))
registerDoParallel(cores=10)
options(warn=0)

## Required packages for saving results
suppressPackageStartupMessages(require(openxlsx))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(cowplot))
suppressPackageStartupMessages(library(Seurat))
#these functions are from Sam Kazar's 2020 Nature Med paper with minor modifications and added notes
{# Choosing the appropriate power for generating the adjacency matrix.
  FindPower <- function(datExpr){
    #choose soft-threshold power
    powers = c(c(1:10), seq(from = 12, to=20, by=2))
    sft=pickSoftThreshold(datExpr,dataIsExpr = TRUE,powerVector = powers,corFnc = cor,
                          corOptions = list(use = 'p', method = "pearson"),networkType = "signed")
    
    # Plot the results
    par(mfrow = c(1,2));
    cex1 = 0.9;
    
    # Scale-free topology fit index as a function of the soft-thresholding power
    plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit, signed R^2",type="n", main = paste("Scale independence"));
    text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],labels=powers,cex=cex1,col="red");
    
    # Red line corresponds to using an R^2 cut-off
    abline(h=0.80,col="red")
    
    # Mean connectivity as a function of the soft-thresholding power
    plot(sft$fitIndices[,1], sft$fitIndices[,5],xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",main = paste("Mean connectivity"))
    text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
  }
  
  # Generating the adjacency matrix and performing clustering
  ClusterTOM <- function(datExpr, softPower){
    #dev.off()
    #Calclute the adjacency matrix
    adj= adjacency(datExpr,type = "signed", power = softPower);
    
    #Turn adjacency matrix into a topological overlap matrix (TOM) to minimize the effects of noise and spurious associations.
    TOM=TOMsimilarityFromExpr(datExpr,networkType = "signed", TOMType = "signed", power = softPower, corType="bicor");
    
    colnames(TOM) = rownames(TOM) = colnames(datExpr)
    dissTOM=1-TOM
    
    #Hierarchical clustering of the genes based on the TOM dissimilarity measure
    geneTree = flashClust(as.dist(dissTOM),method="complete");
    
    #Plot the resulting clustering tree (dendrogram)
    plot(geneTree, xlab="", sub="",cex=0.3);
    
    return(list(dissTOM = dissTOM, geneTree = geneTree)) #returns list with dissimilarity TOM, and the clustered gene tree.
  }
  
  # Cut the resulting clustering dendrogram using the "tree" method for cutreeDynamic. Minimum module size can be specified.
  CutTOMTree <- function(datExpr, dissTOM, geneTree, minModuleSize = 10){
    #dev.off()
    # Module identification using dynamic tree cut, you can also choose the hybrid method
    dynamicMods = cutreeDynamic(dendro = geneTree,  method="tree", minClusterSize = minModuleSize);
    #dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM, method="hybrid", deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = minModuleSize);
    
    #Get the module labels and the size of each module. Lable 0 is reserved for unassigned genes
    print(table(dynamicMods))
    
    #Plot the module assignment under the dendrogram; note: The grey color is reserved for unassigned genes
    dynamicColors = labels2colors(dynamicMods)
    table(dynamicColors)
    plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut", dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05, main = "Gene dendrogram and module colors")
    
    #Set the diagonal of the dissimilarity to NA 
    diag(dissTOM) = NA;
    
    #extract modules
    module_colors= setdiff(unique(dynamicColors), "grey")
    modules = lapply(module_colors, function(x){colnames(datExpr)[which(dynamicColors==x)]})
    names(modules) = module_colors
    return(list(dyanmicColors = dynamicColors, modules = modules)) #returns list with module colors, and the modules themselves
  }
  
  # Merge modules with low dissimilarity. Cutoff for dissimilarity merge can be specified
  MergeSimilarModules <- function(datExpr, dynamicColors, geneTree, MEDissThres = 0.5){
    #cacluate eigengenes
    MEList = moduleEigengenes(datExpr, colors=dynamicColors)
    MEs = MEList$eigengenes
    
    # Calculate dissimilarity of module eigengenes
    MEDiss = 1-cor(MEs);
    
    # Cluster module eigengenes
    METree = hclust(as.dist(MEDiss), method = "average");
    
    # Plot the result
    #sizeGrWindow(7, 6)
    plot(METree, main = "Clustering of module eigengenes",
         xlab = "", sub = "")
    abline(h = MEDissThres, lwd=2, col="red")
    
    # Call an automatic merging function
    merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)
    # The merged module colors
    mergedColors = merge$colors;
    # Eigengenes of the new merged modules:
    mergedMEs = merge$newMEs;
    #plot showing how merged modules exist
    plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                        c("Dynamic Tree Cut", "Merged dynamic"),
                        dendroLabels = FALSE, hang = 0.03,
                        addGuide = TRUE, guideHang = 0.05)
    
    #extract merged modules
    merged_module_colors= setdiff(unique(mergedColors), "grey")
    merged_modules = lapply(merged_module_colors, function(x){colnames(datExpr)[which(mergedColors==x)]})
    names(merged_modules) = merged_module_colors
    
    return(list(mergedColors = mergedColors, merged_modules = merged_modules)) #returns list with merged colors, and the merged modules themselves
  }
  
  ############################################################################################################
  ## Functions for assaying module significance against background and temporal variability in module score ##
  ############################################################################################################
  
  ## Test to determine if the genes within the module are truly the least dissimilar compared to randomly generated modules of the same size.
  TestModuleSignificance <- function(mod, dissTOM, expr.data, n_perm = 10000, pval = 0.05, n.bin = 10, seed = 42){
    #vectorize the actual distribution of (dis)similarities, and remove zeros!
    set.seed(seed)
    true.diss = as.vector(dissTOM[mod,mod])
    true.diss = true.diss[-which(true.diss == 0)]
    
    #size of module for permutations
    mod.size = length(mod)
    
    #bin all genes by expression
    expr.avg = rowMeans(expr.data)
    expr.avg = expr.avg[order(expr.avg)]
    expr.avg.cut = as.numeric(x = cut2(x = expr.avg, m=round(length(expr.avg)/n.bin)))
    names(expr.avg.cut) = names(expr.avg)
    
    #create a table of binnings of all genes and of our module genes
    all.bin.table = table(expr.avg.cut)
    mod.bin.table = table(expr.avg.cut[mod])
    
    #randomly generate module with same expression binning structure and run t.test and record results
    test.results = data.frame(statistic = rep(NA, n_perm), pval = rep(NA, n_perm)) #container for results
    
    for (i in 1:n_perm){ #by permutation
      random.mod = list() #create an empty list we will fill with gene names (each element will be gene names by bin)
      
      #for each bin of the mod bin table, randomly select that number of genes from the full set with that expression
      for (j in names(mod.bin.table)){ 
        bin.genes = sample(names(expr.avg.cut)[which(expr.avg.cut == as.numeric(j))], mod.bin.table[j], replace = FALSE)
        random.mod[[as.numeric(j)]] = bin.genes #stick those genes into the random.mod list
      }
      #unlist and vectorize the distribution of (dis)similarities (remember to remove zeros)
      random.mod = unlist(random.mod)
      random.diss = as.vector(dissTOM[random.mod,random.mod])
      random.diss = random.diss[-which(random.diss == 0)]
      
      #perform a one-sided wilcox.test and record the statistic and p-value.
      #Note, IMPORTANT: here we perform the test asking if the true diss is LESS THAN the random diss, as we are trying to minimize dissimilarity
      test = wilcox.test(x = true.diss, y = random.diss, alternative = "less")
      test.results[i,] = c(test$statistic, test$p.value)
    }
    
    #correct for multiple hypothesis testing, and then report the proportion of bad tests
    test.results$FDR = p.adjust(test.results$pval, method = "fdr")
    num.failed.tests = sum(test.results$FDR > pval)
    print(paste(paste(num.failed.tests, n_perm, sep="/"), "permutations failed the Mann-Whitney test.", sep=" "))
    
    #is the percentage of failed tests less than or equal to the p-val specified?
    return(num.failed.tests/n_perm <= pval) #returns a vector of booleans indicating if each module was significant based on the specific p.val
  }
  
 
}


#This function just wraps the above with appropriate prompts to identify modules 
#By default this is using the "RNA" assay. Modifying it is trivial but 
#default is set to protect against incorrectly using SCT 
ModID<-function(seuratobj, cluster=NA, prefix="modules", nPCS="", softpower=""){
  #subset to a cluster or clusters of interest if desired 
  if (sum(!is.na(cluster))>1){test_clus<-subset(seuratobj, idents=cluster)}else{test_clus<-seuratobj}
  #perform feature selection, scaling, and PCA 
  test_clus<-FindVariableFeatures(test_clus, verbose=FALSE)
  test_clus<-ScaleData(test_clus, verbose=FALSE)
  test_clus<-RunPCA(test_clus, verbose=FALSE)
  
  #prompot for number of PCs needed 
  if(nPCS==""){print(ElbowPlot(test_clus))}
  while(nPCS==""){nPCS<-readline("how many NPCs?")}
  nPCS<-as.integer(nPCS)
  #expanding the genes by projecting the dims to more PCs 
  test_clus<-ProjectDim(test_clus)
  
  #gere we'll grab the top 25 features (positive and negative) associated with each dimension 
  genes<-c()
  for (i in 1:nPCS){genelist<-TopFeatures(test_clus, dim = i, nfeatures = 50,balanced = T, projected = T )
  for (i in 1:length(genelist)){genes<-c(genes, genelist[[i]])}}
  #remove duplicates 
  genes<-unique(genes)
  #grab the matrix from the seurat object 
  test_clus<-as.matrix(test_clus@assays$RNA@data[genes,])
  #prompt for softpower 
  FindPower(datExpr=t(test_clus))
  while(softpower==""){softpower<-readline("what softpower?")}
  softpower<-as.integer(softpower)
  #run the remaining functions to actually cut the tree
  test_clus_tom<-ClusterTOM(datExpr = t(test_clus), softPower = softpower)
  test_clus_mods<-CutTOMTree(datExpr = t(test_clus), geneTree = test_clus_tom$geneTree, dissTOM = test_clus_tom$dissTOM, minModuleSize = 10 )
  test_clus_merge_mods<-MergeSimilarModules(datExpr=t(test_clus), dynamicColors = test_clus_mods$dyanmicColors, geneTree = test_clus_tom$geneTree, MEDissThres = 0.5)
  print(test_clus_merge_mods$merged_modules)                                          
  test_clus_merge_mods.isSig = sapply(test_clus_merge_mods$merged_modules, function(module){
    TestModuleSignificance(mod = module, dissTOM = test_clus_tom$dissTOM, expr.data = test_clus,
                           n_perm = 10000, pval = 0.05, n.bin = 10)
  })
  test_clus_merge_mods.isSig = test_clus_merge_mods$merged_modules[test_clus_merge_mods.isSig]
  print(test_clus_merge_mods.isSig)
  names(test_clus_merge_mods.isSig)<-paste(prefix,names(test_clus_merge_mods.isSig ), sep="_")
  return(test_clus_merge_mods.isSig)
}

#lets me wrap a whole group of objects
WarpGroupModID<-function(seuratobjlist, cluster, conds=""){mods<-list()
for (i in 1:length(seuratobjlist)){
  
  print(names(seuratobjlist)[[i]])
  if(length(conds)==1){res<-try(ModID(seuratobjlist[[i]],cluster=cluster, prefix=names(seuratobjlist)[[i]]))}else{
    res<-try(ModID(seuratobjlist[[i]],cluster=cluster, prefix=names(seuratobjlist)[[i]], nPCS =conds[i,1], softpower=conds[i,2]))
  }
  if(inherits(res, "try-error"))
  {
    #error handling code,  just skip this iteration if not modules get found or theres some error in the process
    next
  }
  
  
  mods[[i]]<-res}
return(mods)
}
#this function clusters mods and allows the user to select a cutoff for merging
MergeMods<-function(modslist,height=""){
  #first find the overlap bettween modules 
  modslist<-modslist[!unlist(lapply(modslist, is.null))]
  name<-c()
  for(i in 1:length(modslist)){
    name<-c(name, names(modslist[[i]]))
  }
  modslist<-unlist(modslist, recursive = FALSE)
  overlaps<-list()
  for(i in 1:length(modslist)){
    modoverlap<-list()
    for(j in 1:length(modslist)){
      modoverlap[[j]]<-length(intersect(modslist[[i]], modslist[[j]]))/length(modslist[[i]])

    }
    overlaps[[i]]<-modoverlap
  }
  
  #datawrangle that for plotting
  data<-matrix(unlist(overlaps), nrow = length(overlaps))
  rownames(data)<-colnames(data)<-name
  p1<-heatmap(data, keep.dendro=TRUE)
  print(heatmap(data))
  res2 = as.hclust( p1$Rowv )
  print(plot(res2))
  #ask for user cut 
  if(height==""){goodenough<-FALSE
  while(goodenough==FALSE){height<-readline(prompt ="where to make the cut")
  if(height==""){height=2}
  res<-cutree(res2, h=as.numeric(height))
  print("number of modules is")
  print(max(res))
  print(plot(res2))
  rect.hclust(res2, h=as.numeric(height), border=2:6)
  goodenough<-readline(prompt = "(w)rite mods or (c)ut again?")
  goodenough<-case_when(goodenough=="w"|goodenough=="W" ~TRUE, TRUE ~FALSE)
  }}else{
      res<-cutree(res2, h=as.numeric(height))
      rect.hclust(res2, h=as.numeric(height), border=2:6)
  }
  #merge based on those identified clusters
  mergedmods<-list()
  for( i in 1:max(res)){
    mergedmods[[i]]<-unique(unlist(modslist[res==i]))
  }
  #return modules 
  return(mergedmods)}



#loading data
setwd("~/gibbs/DOGMAMORPH/Ranalysis")
results<-readRDS( "Objects/20230518Completeobj.rds")
#setting correct default
DefaultAssay(results)<-"RNA"

```

#Mod identification

Mod id ends up being art as much as it is science and is overall an iterative process. The principal goal of mod detection is to identify correlated networks of genes that are presumably corregulated. You can then annotate those modules based on their predicted function and association with disease state. In practice, you can identify almost unlimited number of modules based on the question you ask and the way you decide to break up groups of cells. 

To illustrate this, consider the conditions we have in our dataset: 
treatment (1 of 3 options), timepoint (1 of 2 options), cluster/cell type (1 of 5-15 options), participant (1 of 12), which lead to dozens of modules just on one variable, with lots more possible if you combine conditions. It's at this point that you need to focus on specific questions you want to determine. 

##Mods across all data 
Here I search for modules across the entire dataset. I select a softpower of 7 (the first with a cor > 0.8, the red line) and 6 PCs (where the "elbow" lies). We initially find 5 modules, 2 of which merge into 1 resulting in 4 to test. From those 1 fails the permutation test, and we have three modules. 

Based on experience, I suspect one module corresponds to Myleoid based on the FLT3 and CST3, one is proliferation based on MKI67 and TOP2A, and the final one is B cells based on MS4A1. 

In this case, none are particularly interesting on their own. We may be able to see if this is different across individuals, but it's somewhat unlikely they are helpful here. 

```{r single mod cluster example, fig.width= 16, fig.height=9}
mods<-ModID(results, nPCS = 6, softpower = 7)

```
## Mods in specific cell subsets  
It's somewhat more informative to ask about specific cell types. For example, what are the modules that are within B cells (or B like cells) in the dataset. From the initial annotation, I know cluster 5 and cluster 10 contain the B like cells, so I pass those to subset. 

I ended up initially selecting 8 npcs, and 8 softpower by the same criteria above. In this case, that identified only two modules. It's up to your preference how many you get, but generally speaking 3 is a good objective. You can typically get more modules by reducing the softpower, 1 at a time. These modules will be larger, but the significance should be maintained. You can also decrease the number of npcs but this is unlikely to impact the module recovery at higher softpowers. To demo, there are also corresponding plots for softpower 7, (npcs 8) and softpower 8 (npcs 5)
```{r b cell modules, fig.width= 16, fig.height=9}

mods<-ModID(results, cluster = c(5,10), nPCS = 8, softpower = 8)
```

```{r b cell modules2, fig.width= 16, fig.height=9}

mods<-ModID(results, cluster = c(5,10), nPCS = 8, softpower = 7)
```

```{r b cell modules3, fig.width= 16, fig.height=9}

mods<-ModID(results, cluster = c(5,10), nPCS = 5, softpower = 8)
```

## Module identification across individuals 

In some cases it is desirable to detect the modules across individuals because individual to individual variability may drive interesting variation. In this case, we look at some CD4 clusters across individuals. When getting to these more targetted questions, we need to be careful to ensure there are "enough" cells to make inferences. Anecdotally, this is around ~50 cells, which can be limiting in rare populations or subsets. 

We could do this one by one, but I tend to use the WrapGroupModID function, grabbing the clusters associated with Memory CD4s in the previous annotation - Clusters 1:4

Each participant will get a separate softpower and npcs. This is fine, but you should investigate any large discrepancies. Npcs besides 3-10 or softpowers outside of +/-2 the mean are odd. For reproducibility, these are all captured here. 

Practically speaking, softpowers >9 rarely result in usable modules. If a softpower peaks above 0.8 at softpower >9, I generally select the softpower with the greatest coefficient <9. 

Participant    NPCs    Softpower   Mods   notes
* 5018    6    6    6
* 5014    8    7    5  
* 5017    6    5    5
* 5023    5    4    5       note: Softpower resulted in nothing close to 0.8
* 5009    5    8    3
* 5021    5    9    2       note: Softpower resulted in nothing close to 0.8
* 5037    6    9    0       note: Softpower resulted in nothing close to 0.8
* 5003    3    9    0
* 5016    6    8    3
* 5020    6    8    2
* 5035    6    7    3
* 5026    6    5    4

In this case the mean softpower is 6, so the 9 softpowers are outisde the expected range. It might suggest another round should be run. 

This will result in a list of modules, 1 per Seurat object in the list. In this case, there are a total of 38 modules across 10 individuals. This is sort of unwieldy because annotating them at that point will be difficult. In my experience, there is usually also high redundancy between modules. I usually take strategies to then merge the mods. 

In the Immunity paper, we used MergeMods which performs heirarchical clustering on the module overlap to merge into a particular set that is a union of all modules. 

A second method, programmed below, can instead try to identify "core" modules based on coocurrance of genes across modules. 

First method, ended up with a cut of 1.5 - resulting in 9 mods
```{r modid CD4 across people, fig.width= 16, fig.height=9}
conds<-cbind(c(6,8,6,5,5,5,6,3,6,6,6,6),c(6,7,5,4,8,9,9,9,8,8,7,5))
results_participant_split<-SplitObject(results, split.by = "Participant")
mods<-WarpGroupModID(results_participant_split, c(1,2,3,4), conds=conds)
saveRDS(mods, "~/gibbs/DOGMAMORPH/Ranalysis/modules/MemoryCD4Participant.rds")

consensusmods<-MergeMods(mods, height = 1.5)
saveRDS(consensusmods, "~/gibbs/DOGMAMORPH/Ranalysis/modules/MemoryCD4Participant_consensus.rds")

```

### rationale for merging other ways, in a further doc I'll look at differences in their specificity/batch 

The previous method has the goal of capturing the total variability across individuals. But, that may be very noisy, capturing a lot of genes that are not related. This method looks to identify genes that have a core more set. In this case I'm going step by step to explain, but in practice this can just be used as the below function. 
```{r merging based on comembership, fig.width= 16, fig.height=9}
#change into just a list of modules not a list of lists of modules 
mods_flat<-unlist(mods, recursive = FALSE)
#now we get a list of all unique genes across all modules 
genes<-unique(unlist(mods_flat))
#then we need to calculate a matrix of gene x gene to get the percentage of modules that gene occurs in that another gene also occurs in 

#now we'll iterate through genes to find joint membership across modules for each 
#this step is computationally expensive unless we first make this lookup table
#lookup table just has each gene on the column and each module on the rows 
# a row has a 1 if the gene is present in the module, 0 otherwise 
lookup<-data.frame(matrix(0, nrow = length(mods_flat), ncol=length(genes)))
colnames(lookup)<-genes
for(i in genes){
  for(j in 1:length(mods_flat)){
    lookup[j,i]<-sum(i %in% mods_flat[[j]])
  }
}
#now we can use that lookup to generate results 
#first we make a 0 matrix. 
results_frame<-data.frame(matrix(0, nrow = length(genes), ncol=length(genes)))
colnames(results_frame)<-rownames(results_frame)<-genes

#then iterate through and find what proportion of modules each gene coocurs in 
#it's important that its divided by the total # of modules to reduce influence of singletongene-module pairs. 
for (i in genes){
  for (j in genes){
    results_frame[i,j]<-sum(lookup[[i]]&lookup[[j]])/nrow(lookup)
  }
}

#then we can querry specific genes of interest, for example tox 
sort(results_frame["TOX",], decreasing=TRUE)
#most genes never occur with tox, but some subset are greater than 50% of modules, which may suggest a core module subset 
sort(results_frame["ZEB2",], decreasing=TRUE)
sort(results_frame["CD28",], decreasing=TRUE)
sort(results_frame["CCR6",], decreasing=TRUE)

#we can also use heirarchical clustering to see if there are patterns here to pickup on. 
#we can then make another cut. The difference here is we'll pick something that results in many singletons that can be fitlered out 

  data<-matrix(unlist(results_frame), nrow = length(results_frame))
  rownames(data)<-colnames(data)<-rownames(results_frame)
  p1<-heatmap(data, keep.dendro=TRUE)
  print(heatmap(data))
  res2 = as.hclust( p1$Rowv )
  print(plot(res2))
  abline(0.55, 0)
  height=0.55
  
  res<-cutree(res2, h=as.numeric(height))
  print("number of modules is")
  print(max(res))
  print(plot(res2))
  rect.hclust(res2, h=as.numeric(height), border=2:6)
  
  #merge based on those identified clusters
  mergedmods<-list()
  for( i in 1:max(res)){
    mergedmods[[i]]<-names(res)[res==i]
    }
  
#now we want to filter out this new module set to remove modules that are frequently singletons 

singletonlist<-colnames(lookup)[colSums(lookup)==1]
consensusmods2<-list()
for (i in mergedmods){
  print(sum(i%in%singletonlist)/length(i))
  if((sum(i%in%singletonlist)/length(i))<0.3){consensusmods2[length(consensusmods2)+1]<-list(i)}
}

# in this case we can see that 2 modules have >30% singletons, while the rest occur across individuals, representing a better "core" 
#4 modules remain
saveRDS(consensusmods2, "~/gibbs/DOGMAMORPH/Ranalysis/modules/CD4memory_core.rds")
#as a final iteration, we can combine the two methods, getting a union set from the first method and then filtering out singletons from each of those modules separately 
consensusmods3<-list()
for (i in consensusmods){
  consensusmods3[length(consensusmods3)+1]<-list(i[!i%in%singletonlist])
}
saveRDS(consensusmods3, "~/gibbs/DOGMAMORPH/Ranalysis/modules/CD4memory_consensus_no_singletons.rds")


```


## Downstream 

From here the annotation is up to you and what makes sense for your question. These are at the end of the day gene lists and they can be scored as such via GO/IPA/ENRICHr/Etc. 
You can also send this back into the single cell object and score each module as below with the consensus mods and visualize them, keep in mind your initial population and question when inturpretting the results. IE if we get B cell annotated modules in our T cell results, something has wrong.

```{r module scoring, fig.width= 16, fig.height=9}
results<-AddModuleScore(results, consensusmods3)
FeaturePlot(results, features = paste0("Cluster", 1:9), ncol=4, min.cutoff = 'q5', max.cutoff = 'q95')

```

Depending on covariates you might also want to check if any are associated with a particular condition 
In this case we check participants, condition, and timepoint. 

No module is dominated by just one participant or just one treatment, but it does appear 4 and 6 are slightly elevated in timepoint 3months
```{r module covariates, fig.width= 16, fig.height=9}
VlnPlot(results, features = paste0("Cluster", 1:9),group.by = "Participant", ncol=4,pt.size = 0)
VlnPlot(results, features = paste0("Cluster", 1:9),group.by = "Timepoint", ncol=4,pt.size = 0)
VlnPlot(results, features = paste0("Cluster", 1:9),group.by = "Treatment", ncol=4,pt.size = 0)

```

We can also look at combinations, such as timepoint and treatment 
We can then see interesting patterns of modules that are associated across our conditions, especially if we focus in on the clusters we initially identified these from
Ex module 3 and 9 both appear down in 3 months treatment Naltraxone but not the other two
```{r module combo covariates, fig.width= 16, fig.height=9}
results$TP_Treatment<-paste(results$Treatment, results$Timepoint)
VlnPlot(results, features = paste0("Cluster", 1:9),group.by = "TP_Treatment", ncol=4,pt.size = 0)
VlnPlot(results, features = paste0("Cluster", 1:9),group.by = "TP_Treatment",idents = c(1,2,3,4), ncol=4,pt.size = 0)


```


```{r session info}
devtools::session_info()
```