---
title: "Cytotoxic T high resolution analysis"
author: "Jack Collora"
date: "2023-06-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis objectives 

Based on the previous results in the chromvar, DE testing, and proportion analysis, it seems reasonable to focus on these cells. The specific quesitons to answer are:

1. What are the transcription factor activities that vary across conditions  
2. What are the proteins different across conditions  
3. What can we conclude about the transcriptional profiles different across individuals by  
 * GSEA  
 * GO analysis  
 
```{r loading in data libraries and GSEA functions}
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(fgsea))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(Seurat))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(msigdbr))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(Signac))
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(BiocParallel))

register(MulticoreParam(4))

GSEA<-function(findmarks, genesets){
  findmarks$gene<-rownames(findmarks)
  genes<-findmarks %>%
    arrange(desc(avg_log2FC)) %>% 
    dplyr::select(gene, avg_log2FC)
  rnk<- deframe(genes)
  genes<- deframe(genes)
  genes<- fgsea(pathways=genesets, stats = genes,eps=FALSE)
  genes <- genes %>%
    as_tibble() %>%
    arrange(desc(NES))
  genes<-list(genes, rnk)
  names(genes)<-c("results","rnk")
  return(genes)
}

#Function augments the existing result table with pathway information, including the entire rank order list of gene hits and NES scores 
GSEATable<-function(GSEAwrap_out,gmt, gseaParam=1, name){
  #doing constants first since those are easy
  finalresults<-GSEAwrap_out[1][[1]]
  #first get the order of genes for each pathway, modified from the fGSEA plot function  
  rnk <- rank(GSEAwrap_out[2][[1]])
  ord <- order(rnk, decreasing = T)
  statsAdj <- GSEAwrap_out[2][[1]][ord]
  statsAdj <- sign(statsAdj) * (abs(statsAdj) ^ gseaParam)
  statsAdj <- statsAdj / max(abs(statsAdj))
  pathwaysetup<-list()
  NES<-list()
  print("start ranking")
  #basically here I just want to get everything into a character vector
  for (i in 1:length(finalresults$pathway)){
    pathway <- unname(as.vector(na.omit(match(gmt[[finalresults$pathway[[i]]]], names(statsAdj)))))
    pathway <- sort(pathway)
    NES[[i]]<-calcGseaStat(statsAdj, selectedStats = pathway,returnAllExtremes = TRUE)
    NES[[i]]<-NES[[i]]$tops
    pathwaysetup[[i]]<-pathway}
  print("done ranking")
  finalresults$rnkorder<-pathwaysetup
  finalresults$NESorder<-NES
  finalresults$ID<-name
  return(finalresults)
}


GSEAbig<-function(listofGSEAtables){
  GSEA<-bind_rows(listofGSEAtables)
  return(GSEA)
}



GSEAEnrichmentPlotComparison<-function(PathwayName, GSEACompOut, returnplot="none", cols=NA){
  top<-max(unlist(GSEACompOut$rnkorder))
  GSEACompOut<-GSEACompOut[GSEACompOut$pathway==PathwayName,]
  curves<-list()
  for (i in 1:length(rownames(GSEACompOut))){curves[[i]]<-list(c(0,GSEACompOut$rnkorder[[i]],top),c(0,GSEACompOut$NESorder[[i]],0))
  curves[[i]]<-as.data.frame(curves[[i]])
  curves[[i]]$name<-GSEACompOut$ID[[i]]
  names(curves[[i]])<-c("Rank","ES","Name")}
  curves<-bind_rows(curves)
  if(returnplot=="ES"){p<-EnrichmentScorePlot(curves)
  return(p)}
  if(returnplot=="BC"){p<-BarcodePlot(curves)
  return(p)}
  if(returnplot=="BOTH"){p<-ComboPlot(curves, Title = PathwayName, Table=GSEACompOut[GSEACompOut$pathway==PathwayName,c("ID","padj")], cols=cols)
  return(p)}
  return(curves)
}

BarcodePlot<-function(GSEACompPathway,stacked=FALSE){
  if(stacked==TRUE){GSEACompPathway<-split(GSEACompPathway,GSEACompPathway$Name)
  for (i in 1:length(GSEACompPathway)){GSEACompPathway[[i]]$y<-i}
  GSEACompPathway<-bind_rows(GSEACompPathway)
  GSEACompPathway$y<-((-GSEACompPathway$y)+1)
  p<-ggplot(GSEACompPathway,aes(x=Rank, y=y,xend=Rank, yend=y-1,color=Name))+geom_segment()+theme_classic()+theme(axis.title.y=element_blank(),
                                                                                                                  axis.text.y=element_blank(),
                                                                                                                  axis.ticks.y=element_blank())+xlab("Gene Rank") + xlim(-5,max(GSEACompPathway$Rank+100) )
  }else{
    p<-ggplot(GSEACompPathway,aes(x=Rank, y=-1,xend=Rank, yend=1,color=Name))+geom_segment()+theme_classic()
  }
  return(p)}


EnrichmentScorePlot<-function(GSEACompPathway){
  p<-ggplot(GSEACompPathway, aes(x=Rank, y=ES, color=Name))+geom_line(linewidth=1.5)+geom_hline(yintercept = 0)+theme_classic()+theme(axis.title.x=element_blank(),
                                                                                                                                 axis.text.x=element_blank(),
                                                                                                                                 axis.ticks.x=element_blank())+ylab("Enrichment Score")+xlim(-5,max(GSEACompPathway$Rank+100) )
  return(p)}
ComboPlot<-function(GSEACompPathway, Title="Enrichment Plot",Table=NA, cols=NA){
  EP<-EnrichmentScorePlot(GSEACompPathway)
  if(length(Table)>1){EP<-EP+annotation_custom(tableGrob(Table, theme = ttheme_minimal(), rows = NULL, cols = c("Name","q Value")), xmin = 10000, ymin = 0.60)}
  BC<-BarcodePlot(GSEACompPathway, stacked=TRUE)
  if(!is.na(cols[1])){
    BC<-BC+scale_color_manual(values = cols)
    EP<-EP+scale_color_manual(values = cols)
  }
  p<-ggarrange(EP,BC, common.legend = TRUE, ncol = 1, heights = c(0.6,0.25), align = "v", labels = Title)
  
  return(p)}


Meanenrichmentplot<-function(GSEAmain,greplist=c(),pathwaytodo=""){
    final<-list()
    IDssmooth<-paste(greplist, "smooth")
    IDsgray<-paste(greplist, "rough")
    #for each group, go through and make a smooth track 
    for(i in 1:length(greplist)){
    GSEAres<-GSEAmain[grepl(greplist[[i]], GSEAmain$ID) & grepl(pathwaytodo, GSEAmain$pathway),]
    GSEAres$split<-IDsgray[[i]]
    maximum<-max(unlist(GSEAres$rnkorder))
    #generate bins of size 5 (helps with smoothing)
    bins<-seq(1,maximum ,5)
    bins<-c(bins, maximum,maximum)
    res<-list()
    #average anypoint within those bins
    for (j in 1:(length(bins)-1)){
      res[[j]]<-mean(unlist(GSEAres$NESorder)[unlist(GSEAres$rnkorder)>bins[[j]]& unlist(GSEAres$rnkorder)<=bins[[j+1]]])
    }
    res<-c(res, 0)
    #remove any bin that does not have a point 
    bins<-bins[!is.na(res)]
    res<-res[!is.na(res)]
    #make a new track, add it to the overall 
    res<-list(GSEAres$pathway[[1]], NA, NA, NA, NA, NA, NA, NA, list(bins), list(res), "smooth", IDssmooth[[i]])
    GSEAres<-rbind(GSEAres, res)
    GSEAres$NESorder<-lapply(GSEAres$NESorder, as.double)
    curves<-list()
    #generate coordinate curves for each track including the smooth 
    for (j in 1:length(rownames(GSEAres))){curves[[j]]<-list(c(0,GSEAres$rnkorder[[j]],maximum),c(0,GSEAres$NESorder[[j]],0))
    curves[[j]]<-as.data.frame(curves[[j]])
    curves[[j]]$name<-GSEAres$ID[[j]]
    curves[[j]]$split<-GSEAres$split[[j]]
    names(curves[[j]])<-c("Rank","ES","Name","split")}
    final[[i]]<-bind_rows(curves)
    }
    #merge all the groups together
  final<-bind_rows(final)
  #plot things
  final<-ggplot(final[!grepl("smooth", final$split),],aes(x=Rank, y=ES, group=Name,color=split))+geom_line(linewidth=0.25)+geom_hline(yintercept = 0) + 
    geom_smooth(data = final[grepl("smooth", final$split),], aes(x=Rank, y=ES, group=split, color=split), linewidth=1.5, method = "gam")+theme_classic()+
    ggtitle(pathwaytodo)+ylab("Enrichment Score")+xlim(-5,max(maximum+100))
  
  return(final)}
 


GSEATable<-function(GSEAwrap_out,gmt, gseaParam=1, name){
  #doing constants first since those are easy
  finalresults<-GSEAwrap_out[1][[1]]
  #first get the order of genes for each pathway, modified from the fGSEA plot function  
  rnk <- rank(GSEAwrap_out[2][[1]])
  ord <- order(rnk, decreasing = T)
  statsAdj <- GSEAwrap_out[2][[1]][ord]
  statsAdj <- sign(statsAdj) * (abs(statsAdj) ^ gseaParam)
  statsAdj <- statsAdj / max(abs(statsAdj))
  pathwaysetup<-list()
  NES<-list()
  print("start ranking")
  #basically here I just want to get everything into a character vector
  for (i in 1:length(finalresults$pathway)){
    pathway <- unname(as.vector(na.omit(match(gmt[[finalresults$pathway[[i]]]], names(statsAdj)))))
    pathway <- sort(pathway)
    NES[[i]]<-calcGseaStat(statsAdj, selectedStats = pathway,returnAllExtremes = TRUE)
    NES[[i]]<-NES[[i]]$tops
    pathwaysetup[[i]]<-pathway}
  print("done ranking")
  finalresults$rnkorder<-pathwaysetup
  finalresults$NESorder<-NES
  finalresults$ID<-name
  return(finalresults)
}

FixMotifID<-function(markeroutput, seuratobj){
  markeroutput$genes<-ConvertMotifID(seuratobj, assay="ATAC", id=rownames(markeroutput))
  return(markeroutput)
}


GSEAbig<-function(listofGSEAtables){
  GSEA<-bind_rows(listofGSEAtables)
  return(GSEA)
}

VolPlot<-function(results,top=TRUE, adthresh=TRUE, thresh=0.2, threshn=30, Title=NA){
  results$value<-(-log10(results$p_val_adj))
  results$value[results$value==Inf]<-300
  results$gene<-rownames(results)
  p<-ggplot(results,aes(x=avg_log2FC, y=value, label=gene))+geom_point()+theme_classic()+xlab("Average Log Fold Change")+ylab("-log10 adjusted p value")+theme(plot.title = element_text(hjust = 0.5))
  if(adthresh==TRUE){thresh<-AdaptiveThreshold(results, Tstart=thresh,n=threshn)}
  if(top==TRUE){
    p<- p + geom_text_repel(aes(label=ifelse((avg_log2FC>thresh|avg_log2FC<(-thresh))&value>(-log10(0.01)) ,as.character(gene),'')),hjust=1,vjust=1)
  }
  if(!is.na(Title)){
    p<-p+ggtitle(Title)
  }
  return(p)
}
#quick selection of top n genes basically
AdaptiveThreshold<-function(results, n=30,Tstart=0.25){
  m=Inf
  while(n<=m){
    Tstart=Tstart+0.05
    m<-length(rownames(results)[(results$avg_log2FC>Tstart|results$avg_log2FC<(-Tstart))&results$value>10])
  }
  return(Tstart)
}

#borrowing some rushmore colors from the wes anderson color pack
BottleRocket2 = c("#FAD510", "#CB2314", "#273046", "#354823", "#1E1E1E")

#takes in enrichr output
PlotEnrichment<-function(file, topn=10, returnplot=TRUE, title="Significant Pathways"){
  data<-read.delim(file, sep = "\t",header = 1)
  data$value<- -log10(data$Adjusted.P.value)
  #order by most signfiicant 
  data$Term<-factor(data$Term, levels = rev(data$Term))
  data<-data[order(data$value, decreasing = TRUE),]
  #subset to top n 
  data<-data[1:topn,]
  plot<-ggplot(data, aes(x=value, y=Term, fill=ifelse(value>2, "Significant", "Not Significant")))+geom_bar(stat="identity")+xlab("-log10(p_adj)")+ylab("Pathway")+ggtitle(title)+
    theme_classic()+scale_fill_manual(values=c("Significant" = "#CB2314", "Not Significant" = "#1E1E1E"))+ guides(fill=guide_legend(title=""))
  if(returnplot){return(plot)}
  return(data)
}

#modded from WA pallette 
IsleofDogs1<-c("Bup.Nalo_vs_Nal_0" ="#9986A5", "Bup.Nalo_vs_Nal_3" = "#79402E", "Meth_vs_Bup.Nalo_0" = "#CCBA72", 
                 "Meth_vs_Bup.Nalo_3" = "#0F0D0E", "Meth_vs_Nal_0" = "#D9D0D3", "Meth_vs_Nal_3" = "#8D8680")


results<-readRDS("~/gibbs/DOGMAMORPH/Ranalysis/Objects/20230606FinalObj.rds") 

```
 
```{r genesets}

#grabbing hallmark as well as the curated, immune 
m_df_H<- msigdbr(species = "Homo sapiens", category = "H")
m_df_H<- rbind(msigdbr(species = "Homo sapiens", category = "C2"), m_df_H)
m_df_H<- rbind(msigdbr(species = "Homo sapiens", category = "C7"), m_df_H)
fgsea_sets<- m_df_H %>% split(x = .$gene_symbol, f = .$gs_name)
``` 

## transcription factor activities 

The objective is to see if there are specific transcription factors that may be driving differential activity across conditions. We will be comparing at time 0 and time 3 in a nonbias manner (DE) and the graphing specific factors of interest. 

Results suggest that JUN/FOS factors are perturbed in this context, along with a few more housekeeping like features like CTCF and YY1. Further analysis should focus on correlating these factors with gene expression (in the whole dataset, as well as within this population, and across conditions) to see if there is a program driven by this change. 

```{r factor activities across conditions, fig.width=16, fig.height=9}
results$T_Tp<-paste(results$Treatment, results$Timepoint, sep = "_")
results$T_Tp<-factor(results$T_Tp, levels = c("Methadone_0","Methadone_3","Bup.Nalo_0", "Bup.Nalo_3", "Naltrexone_0", "Naltrexone_3"))
results_cyto<-subset(results, merged_clusters=="Cytotoxic_T")

DefaultAssay(results_cyto)<-"chromvar"
Idents(results_cyto)<-results_cyto$T_Tp

# month 3 comparisons 

Meth_vs_Nal_3<-FindMarkers(results_cyto, "Methadone_3","Naltrexone_3", mean.fxn = rowMeans)
Meth_vs_Bup.Nalo_3<-FindMarkers(results_cyto, "Methadone_3","Bup.Nalo_3", mean.fxn = rowMeans)
Bup.Nalo_vs_Nal_3<-FindMarkers(results_cyto, "Bup.Nalo_3","Naltrexone_3", mean.fxn = rowMeans)

# month 0 comparisons 
Meth_vs_Nal_0<-FindMarkers(results_cyto, "Methadone_0","Naltrexone_0", mean.fxn = rowMeans)
Meth_vs_Bup.Nalo_0<-FindMarkers(results_cyto, "Methadone_0","Bup.Nalo_0", mean.fxn = rowMeans)
Bup.Nalo_vs_Nal_0<-FindMarkers(results_cyto, "Bup.Nalo_0","Naltrexone_0", mean.fxn = rowMeans)


Meth_vs_Nal_3<-FixMotifID(Meth_vs_Nal_3, results_cyto)
Meth_vs_Bup.Nalo_3<-FixMotifID(Meth_vs_Bup.Nalo_3, results_cyto)
Bup.Nalo_vs_Nal_3<-FixMotifID(Bup.Nalo_vs_Nal_3, results_cyto)
Meth_vs_Nal_0<-FixMotifID(Meth_vs_Nal_0, results_cyto)
Meth_vs_Bup.Nalo_0<-FixMotifID(Meth_vs_Bup.Nalo_0, results_cyto)
Bup.Nalo_vs_Nal_0<-FixMotifID(Bup.Nalo_vs_Nal_0, results_cyto)

DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Bup.Nalo_vs_Nal_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_0)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_0)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Bup.Nalo_vs_Nal_0)


VlnPlot(results_cyto, "MA1143.1", pt.size = 0.1 )+ggtitle("FOSL1::JUND")
VlnPlot(results_cyto, "MA0095.3", pt.size = 0.1 )+ggtitle("Yy1")
VlnPlot(results_cyto, "MA0139.1", pt.size = 0.1 )+ggtitle("CTCF")
VlnPlot(results_cyto, "MA0656.1", pt.size = 0.1 )+ggtitle("JDP2")
VlnPlot(results_cyto, "MA0605.2", pt.size = 0.1 )+ggtitle("ATF3")
VlnPlot(results_cyto, "MA0840.1", pt.size = 0.1 )+ggtitle("Creb5")
VlnPlot(results_cyto, "MA1145.1", pt.size = 0.1 )+ggtitle("FOSL2::JUND")
VlnPlot(results_cyto, "MA1131.1", pt.size = 0.1 )+ggtitle("FOSL2::JUN")
VlnPlot(results_cyto, "MA1155.1", pt.size = 0.1 )+ggtitle("ZSCAN4")
VlnPlot(results_cyto, "MA0478.1", pt.size = 0.1 )+ggtitle("FOSL2")
VlnPlot(results_cyto, "MA1988.1", pt.size = 0.1 )+ggtitle("Atf3")
VlnPlot(results_cyto, "MA1134.1", pt.size = 0.1 )+ggtitle("FOS::JUNB")
VlnPlot(results_cyto, "MA0489.2", pt.size = 0.1 )+ggtitle("Jun")
VlnPlot(results_cyto, "MA1634.1", pt.size = 0.1 )+ggtitle("BATF")
VlnPlot(results_cyto, "MA0101.1", pt.size = 0.1 )+ggtitle("RELA")


Idents(results_cyto)<-factor(Idents(results_cyto),levels = c("Methadone_0","Bup.Nalo_0","Naltrexone_0","Methadone_3", "Bup.Nalo_3", "Naltrexone_3"))

VlnPlot(results_cyto, "MA1143.1", pt.size = 0.1 )+ggtitle("FOSL1::JUND")
VlnPlot(results_cyto, "MA0095.3", pt.size = 0.1 )+ggtitle("Yy1")
VlnPlot(results_cyto, "MA0139.1", pt.size = 0.1 )+ggtitle("CTCF")
VlnPlot(results_cyto, "MA0656.1", pt.size = 0.1 )+ggtitle("JDP2")
VlnPlot(results_cyto, "MA0605.2", pt.size = 0.1 )+ggtitle("ATF3")
VlnPlot(results_cyto, "MA0840.1", pt.size = 0.1 )+ggtitle("Creb5")
VlnPlot(results_cyto, "MA1145.1", pt.size = 0.1 )+ggtitle("FOSL2::JUND")
VlnPlot(results_cyto, "MA1131.1", pt.size = 0.1 )+ggtitle("FOSL2::JUN")
VlnPlot(results_cyto, "MA1155.1", pt.size = 0.1 )+ggtitle("ZSCAN4")
VlnPlot(results_cyto, "MA0478.1", pt.size = 0.1 )+ggtitle("FOSL2")
VlnPlot(results_cyto, "MA1988.1", pt.size = 0.1 )+ggtitle("Atf3")
VlnPlot(results_cyto, "MA1134.1", pt.size = 0.1 )+ggtitle("FOS::JUNB")
VlnPlot(results_cyto, "MA0489.2", pt.size = 0.1 )+ggtitle("Jun")
VlnPlot(results_cyto, "MA1634.1", pt.size = 0.1 )+ggtitle("BATF")
VlnPlot(results_cyto, "MA0101.1", pt.size = 0.1 )+ggtitle("RELA")

```

## Surface protein differences 

Doing the same thing as Chormvar but instead we're looking at protein expression. 

As below, this cell type basically has no differences in protein expression across conditions. The only significant factor was CD101, which is reported to impact myeloid differentiation. 

```{r proteins across conditions, fig.width=16, fig.height=9}

DefaultAssay(results_cyto)<-"ADT"
Idents(results_cyto)<-results_cyto$T_Tp

#need to first do CLR by library 

results_cyto_split<-SplitObject(results_cyto, split.by = "orig.ident")

results_cyto_split<-lapply(results_cyto_split, NormalizeData, normalization.method = "CLR")

for (i in 1:9){
  results_cyto_split[[i]]<-results_cyto_split[[i]]@assays$ADT@data
}
results_cyto_split<-do.call(cbind, results_cyto_split)
#put it in but ensure things are correct
results_cyto@assays$ADT@data<-results_cyto_split[,colnames(results_cyto)]

Meth_vs_Nal_3<-FindMarkers(results_cyto, "Methadone_3","Naltrexone_3")
Meth_vs_Bup.Nalo_3<-FindMarkers(results_cyto, "Methadone_3","Bup.Nalo_3")
Bup.Nalo_vs_Nal_3<-FindMarkers(results_cyto, "Bup.Nalo_3","Naltrexone_3")

# month 0 comparisons 
Meth_vs_Nal_0<-FindMarkers(results_cyto, "Methadone_0","Naltrexone_0")
Meth_vs_Bup.Nalo_0<-FindMarkers(results_cyto, "Methadone_0","Bup.Nalo_0")
Bup.Nalo_vs_Nal_0<-FindMarkers(results_cyto, "Bup.Nalo_0","Naltrexone_0")

#Bup.Nalo and Nal are not different at all (0 pass logfc)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_0)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_0)

#only DE one, promyeloid differentiation 
VlnPlot(results_cyto, "CD101-TotalA", pt.size = 0.1 )


```

## Gene expression and GSEA 


### Standard DE 

We're just using the standard wilcox/mann whitney U test here to look at the impact of these different conditions on gene expression
```{r standard DE, fig.width=16, fig.height=9}
DefaultAssay(results_cyto)<-"RNA"

# month 3 comparisons 

Meth_vs_Nal_3<-FindMarkers(results_cyto, "Methadone_3","Naltrexone_3")
Meth_vs_Bup.Nalo_3<-FindMarkers(results_cyto, "Methadone_3","Bup.Nalo_3")
Bup.Nalo_vs_Nal_3<-FindMarkers(results_cyto, "Bup.Nalo_3","Naltrexone_3")

# month 0 comparisons 
Meth_vs_Nal_0<-FindMarkers(results_cyto, "Methadone_0","Naltrexone_0")
Meth_vs_Bup.Nalo_0<-FindMarkers(results_cyto, "Methadone_0","Bup.Nalo_0")
Bup.Nalo_vs_Nal_0<-FindMarkers(results_cyto, "Bup.Nalo_0","Naltrexone_0")


DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Bup.Nalo_vs_Nal_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_0)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_0)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Bup.Nalo_vs_Nal_0)

VlnPlot(results_cyto, "KLRG1", pt.size = 0.1)

#volcanos of just those highly differentially expressed
VolPlot(Meth_vs_Nal_3)
VolPlot(Meth_vs_Bup.Nalo_3)
VolPlot(Bup.Nalo_vs_Nal_3)
VolPlot(Meth_vs_Nal_0)
VolPlot(Meth_vs_Bup.Nalo_0)
VolPlot(Bup.Nalo_vs_Nal_0)

#doing it again but with no cutoffs for FC

# month 3 comparisons 

Meth_vs_Nal_3<-FindMarkers(results_cyto, "Methadone_3","Naltrexone_3", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)
Meth_vs_Bup.Nalo_3<-FindMarkers(results_cyto, "Methadone_3","Bup.Nalo_3", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)
Bup.Nalo_vs_Nal_3<-FindMarkers(results_cyto, "Bup.Nalo_3","Naltrexone_3", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)

# month 0 comparisons 
Meth_vs_Nal_0<-FindMarkers(results_cyto, "Methadone_0","Naltrexone_0", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)
Meth_vs_Bup.Nalo_0<-FindMarkers(results_cyto, "Methadone_0","Bup.Nalo_0", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)
Bup.Nalo_vs_Nal_0<-FindMarkers(results_cyto, "Bup.Nalo_0","Naltrexone_0", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)

#not placing data tables for these very large dataframes as they make the resulting html unusable 

VolPlot(Meth_vs_Nal_3)
VolPlot(Meth_vs_Bup.Nalo_3)
VolPlot(Bup.Nalo_vs_Nal_3)
VolPlot(Meth_vs_Nal_0)
VolPlot(Meth_vs_Bup.Nalo_0)
VolPlot(Bup.Nalo_vs_Nal_0)
```
#### GO analysis of DE genes from traditional analyses 

Using Enrichr we can look for genesets that are enriched here for each comparison. 

In general it doesn't look that interesting because the pathways involved are focused mainly on those dominated by ribosomal proteins, although there is a subset that interestingly has IL27 signaling. 

```{r GO_trad, fig.width=16, fig.height=9}
#outputing the genelists to test 

to_output<-list(Meth_vs_Nal_3,Meth_vs_Bup.Nalo_3,Bup.Nalo_vs_Nal_3,Meth_vs_Nal_0,Meth_vs_Bup.Nalo_0,Bup.Nalo_vs_Nal_0 )
names(to_output)<-c("Meth_vs_Nal_3","Meth_vs_Bup.Nalo_3","Bup.Nalo_vs_Nal_3","Meth_vs_Nal_0","Meth_vs_Bup.Nalo_0","Bup.Nalo_vs_Nal_0")
pcut<-0.05 
LFCcut<-0.25
outdir<-"~/gibbs/DOGMAMORPH/Ranalysis/enrichr/Trad_DE_cyto"
#basically we'll loop through, write the subset of genes based on those cutoffs 
for (i in 1:length(to_output)){
  cur<-to_output[[i]][(to_output[[i]]$avg_log2FC< -LFCcut) | (to_output[[i]]$avg_log2FC> LFCcut),]
  cur$p_val_adj<-p.adjust(cur$p_val, "BH")
  write.table(rownames(cur)[(cur$p_val_adj<pcut)&(cur$avg_log2FC>0)], paste0(outdir,"/", names(to_output)[[i]], "_up.txt"), 
              quote = FALSE, row.names = FALSE, col.names = FALSE)
  write.table(rownames(cur)[(cur$p_val_adj<pcut)&(cur$avg_log2FC<0)], paste0(outdir,"/", names(to_output)[[i]], "_down.txt"), 
              quote = FALSE, row.names = FALSE, col.names = FALSE)
}


todo<-list.files("~/gibbs/DOGMAMORPH/Ranalysis/enrichr/Trad_DE_cyto/Results_tables/")

for (i in todo){
  print(PlotEnrichment(paste0("~/gibbs/DOGMAMORPH/Ranalysis/enrichr/Trad_DE_cyto/Results_tables/",i), topn = 20, title=i))
}

```
#### GSEA analysis of DE genes from traditional analyses 

GSEA is a nice addition particularly in single cell because the signals are weaker due to inherent sparsity and pooling info across genes by ranking can be more powerful. 

```{r GSEA_trad, fig.width=16, fig.height=9}
GSEAres<-list()
for (i in 1:length(to_output)){
GSEAres[[i]]<-GSEA(to_output[[i]], genesets = fgsea_sets)
GSEAres[[i]]<-GSEATable(GSEAwrap_out =GSEAres[[i]], gmt = fgsea_sets, name = names(to_output)[[i]] )}
GSEAres<-GSEAbig(listofGSEAtables = GSEAres)

saveRDS(GSEAres, file="~/gibbs/DOGMAMORPH/Ranalysis/Objects/20230622GSEACtyoTTrad.rds")
#identified these in 20230627cytotoxicGSEA.R

to_test<-c("PHONG_TNF_TARGETS_UP", "GSE18791_CTRL_VS_NEWCASTLE_VIRUS_DC_6H_DN","GSE6269_FLU_VS_E_COLI_INF_PBMC_DN",
           "KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION", "BOSCO_INTERFERON_INDUCED_ANTIVIRAL_MODULE", "GSE34205_HEALTHY_VS_RSV_INF_INFANT_PBMC_UP", 
           "REACTOME_CYTOKINE_SIGNALING_IN_IMMUNE_SYSTEM", "GSE22886_NAIVE_CD8_TCELL_VS_MEMORY_TCELL_UP",
           "DANG_BOUND_BY_MYC","GSE34205_HEALTHY_VS_FLU_INF_INFANT_PBMC_UP","MARSON_BOUND_BY_FOXP3_UNSTIMULATED","ZHENG_BOUND_BY_FOXP3",
           "GSE26030_TH1_VS_TH17_RESTIMULATED_DAY5_POST_POLARIZATION_UP", "ICHIBA_GRAFT_VERSUS_HOST_DISEASE_D7_UP",
           "GSE21670_UNTREATED_VS_TGFB_IL6_TREATED_CD4_TCELL_DN", "WP_ELECTRON_TRANSPORT_CHAIN_OXPHOS_SYSTEM_IN_MITOCHONDRIA",
           "GSE24574_BCL6_HIGH_TFH_VS_TFH_CD4_TCELL_DN","GSE2770_TGFB_AND_IL4_ACT_VS_ACT_CD4_TCELL_2H_UP","HALLMARK_TNFA_SIGNALING_VIA_NFKB",
           "GSE21063_WT_VS_NFATC1_KO_16H_ANTI_IGM_STIM_BCELL_UP","GSE41978_KLRG1_HIGH_VS_LOW_EFFECTOR_CD8_TCELL_DN",
           "ICHIBA_GRAFT_VERSUS_HOST_DISEASE_35D_UP", "BLANCO_MELO_RESPIRATORY_SYNCYTIAL_VIRUS_INFECTION_A594_CELLS_UP",
           "BROWNE_HCMV_INFECTION_14HR_UP","WIELAND_UP_BY_HBV_INFECTION","REACTOME_HIV_INFECTION",
           "GSE43955_TH0_VS_TGFB_IL6_IL23_TH17_ACT_CD4_TCELL_60H_DN", "GSE5960_TH1_VS_ANERGIC_TH1_UP", "GSE12392_IFNAR_KO_VS_IFNB_KO_CD8_NEG_SPLEEN_DC_UP")
for (i in 1:length(to_test)){
  print(GSEAEnrichmentPlotComparison(to_test[i], GSEAres, "BOTH", cols = IsleofDogs1))
}

```


### pseudobulk comparisons by timepoint 

Rationale: a few studies now have shown that this is a more robust way to analyze single cell data because it's less susceptible to normalization issues around zero. Basically if we instead of a psuedobulk, although the significance goes down because we are no longer having such an inflated N, we do get more meaningful differences. 

This is most important to to because the principals of many standard statistical tests assume independence, and cells are not independent within a sample. 

Unfortunately, in this dataset this isn't a useful approach because nothing ends up significant by this bar. So we continue with the above where we have many samples as opposed to relatively few. 

```{r DE testing itself}

prepDESeq2<-function(seuratobj, group_col, rep_col){
  expression<-seuratobj@assays$RNA@counts
  results<-list()
  #first aggregate by participant
  for (i in names(table(seuratobj[[rep_col]]))){
    sub<-seuratobj[[rep_col]]==i
    results[[i]]<-expression[,sub[,1]]
    results[[i]]<-rowSums(results[[i]])
  }
  names(results)<-paste0("Psuedobulk_", names(results))
  expression<-as.data.frame(results)
  
  #now create the condition matrix based on that group_col 
  coldata<-group_by(seuratobj@meta.data, !!sym(group_col), !!sym(rep_col))%>%summarise()
  coldata<-as.data.frame(coldata)
  rownames(coldata)<-paste0("Psuedobulk_", coldata[[rep_col]])  
  coldata[,1]<-factor(coldata[,1])
  coldata<-coldata[colnames(expression),]
  #finally, we make a DESeq object
  expression<-DESeqDataSetFromMatrix(countData = expression, colData = coldata,
                         design = as.formula(paste("~ ", group_col) ) )
  return(expression)
}
results_cyto$T_Tp_P<-paste(results_cyto$T_Tp, results_cyto$Participant, sep="_")
dds<-prepDESeq2(results_cyto, "T_Tp","T_Tp_P")
dds <- DESeq(dds)

meth_0_vs_3 <- results(dds, contrast=c("T_Tp", "Methadone_0", "Methadone_3"),alpha=0.05)
meth_0_vs_3 <- lfcShrink(dds, contrast=c("T_Tp", "Methadone_0", "Methadone_3"), res=meth_0_vs_3, type = "ashr")

#no DE genes, so no Enrichr is logical
summary(meth_0_vs_3)

bup.nalo_0_vs_3 <- results(dds, contrast=c("T_Tp", "Bup.Nalo_0", "Bup.Nalo_3"),alpha=0.05)
bup.nalo_0_vs_3 <- lfcShrink(dds, contrast=c("T_Tp", "Bup.Nalo_0", "Bup.Nalo_3"), res=bup.nalo_0_vs_3, type = "ashr")

summary(bup.nalo_0_vs_3)

Nalt_0_vs_3 <- results(dds, contrast=c("T_Tp", "Naltrexone_0", "Naltrexone_3"),alpha=0.05)
Nalt_0_vs_3 <- lfcShrink(dds, contrast=c("T_Tp", "Naltrexone_0", "Naltrexone_3"), res=Nalt_0_vs_3, type = "ashr")

summary(Nalt_0_vs_3)

Meth_vs_Nal_3<- results(dds, contrast=c("T_Tp", "Methadone_3", "Naltrexone_3"), alpha=0.05)
Meth_vs_Bup.Nalo_3<- results(dds, contrast=c("T_Tp", "Methadone_3", "Bup.Nalo_3"), alpha=0.05)
Bup.Nalo_vs_Nal_3<- results(dds, contrast=c("T_Tp", "Bup.Nalo_3", "Naltrexone_3"), alpha=0.05)
Meth_vs_Nal_0<- results(dds, contrast=c("T_Tp", "Methadone_0", "Naltrexone_0"), alpha=0.05)
Meth_vs_Bup.Nalo_0<- results(dds, contrast=c("T_Tp", "Naltrexone_0", "Bup.Nalo_0"), alpha=0.05)
Bup.Nalo_vs_Nal_0<- results(dds, contrast=c("T_Tp", "Methadone_0", "Naltrexone_0"), alpha=0.05)

Meth_vs_Nal_3<- lfcShrink(dds, contrast=c("T_Tp", "Methadone_3", "Naltrexone_3"),
  res=Meth_vs_Nal_3, type = "ashr")
Meth_vs_Bup.Nalo_3<- lfcShrink(dds, contrast=c("T_Tp", "Methadone_3", "Bup.Nalo_3"),
  res=Meth_vs_Bup.Nalo_3, type = "ashr")
Bup.Nalo_vs_Nal_3<- lfcShrink(dds, contrast=c("T_Tp", "Bup.Nalo_3", "Naltrexone_3"),
  res=Bup.Nalo_vs_Nal_3, type = "ashr")
Meth_vs_Nal_0<- lfcShrink(dds, contrast=c("T_Tp", "Methadone_0", "Naltrexone_0"),
  res=Meth_vs_Nal_0, type = "ashr")
Meth_vs_Bup.Nalo_0<- lfcShrink(dds, contrast=c("T_Tp", "Naltrexone_0", "Bup.Nalo_0"),
  res=Meth_vs_Bup.Nalo_0, type = "ashr")
Bup.Nalo_vs_Nal_0<- lfcShrink(dds, contrast=c("T_Tp", "Methadone_0", "Naltrexone_0"),
  res=Bup.Nalo_vs_Nal_0, type = "ashr")


to_output<-list(Meth_vs_Nal_3,Meth_vs_Bup.Nalo_3,Bup.Nalo_vs_Nal_3,Meth_vs_Nal_0,Meth_vs_Bup.Nalo_0,Bup.Nalo_vs_Nal_0 )
names(to_output)<-c("Meth_vs_Nal_3","Meth_vs_Bup.Nalo_3","Bup.Nalo_vs_Nal_3","Meth_vs_Nal_0","Meth_vs_Bup.Nalo_0","Bup.Nalo_vs_Nal_0")
lapply(to_output, summary)

```


```{r session info}
devtools::session_info()
```