---
title: "CytotoxicTHighRes_ATAC"
author: "Jack Collora"
date: "2023-06-27"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis objectives 
Determine if there are ATAC-seq differences across conditions and timepoints. 

Spoiler there isn't. 


```{r loading}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(Seurat))
suppressPackageStartupMessages(library(Signac))

results<-readRDS("~/gibbs/DOGMAMORPH/Ranalysis/Objects/20230606FinalObj.rds") 
DefaultAssay(results)<-"ATAC"
results$T_Tp<-paste(results$Treatment, results$Timepoint, sep = "_")
results$T_Tp<-factor(results$T_Tp, levels = c("Methadone_0","Methadone_3","Bup.Nalo_0", "Bup.Nalo_3", "Naltrexone_0", "Naltrexone_3"))
results_cyto<-subset(results, merged_clusters=="Cytotoxic_T")
Idents(results_cyto)<-results_cyto$T_Tp

```

## Differential accessibility 

Goal here is to look for peaks that are different across conditions to see if we can find epigenetic changes/priming differences that are relevant in these treatments 

### same condition across timepoint 

First we'll try each condition across time. There are none different across. 


```{r methadone Bup.Nalo Nalt}
#none significant in any comp
Methadone_3_vs_0 <- FindMarkers(
  object = results_cyto,
  ident.1 = 'Methadone_3',
  ident.2 = 'Methadone_0',
  only.pos = TRUE,
  test.use = 'LR',
  min.pct = 0.05,
  latent.vars = 'nCount_ATAC'
)

Bup.Nalo_3_vs_0 <- FindMarkers(
  object = results_cyto,
  ident.1 = 'Bup.Nalo_3',
  ident.2 = 'Bup.Nalo_0',
  only.pos = TRUE,
  test.use = 'LR',
  min.pct = 0.05,
  latent.vars = 'nCount_ATAC'
)

Nalt_3_vs_0 <- FindMarkers(
  object = results_cyto,
  ident.1 = 'Naltrexone_3',
  ident.2 = 'Naltrexone_0',
  only.pos = TRUE,
  test.use = 'LR',
  min.pct = 0.05,
  latent.vars = 'nCount_ATAC'
)

```



### Across conditons at tp 3

Now we're looking at the DE peaks all after treatment, there is 1 different in two comparisons, i.e. no difference. 

```{r methadone Bup.Nalo Nalt each other}
#0 DE peaks
Methadone_3_vs_Bup.Nalo3 <- FindMarkers(
  object = results_cyto,
  ident.1 = 'Methadone_3',
  ident.2 = 'Bup.Nalo_3',
  only.pos = TRUE,
  test.use = 'LR',
  min.pct = 0.05,
  latent.vars = 'nCount_ATAC'
)
#1 DE peak
Bup.Nalo_3_vs_Nalt_3 <- FindMarkers(
  object = results_cyto,
  ident.1 = 'Bup.Nalo_3',
  ident.2 = 'Naltrexone_3',
  only.pos = TRUE,
  test.use = 'LR',
  min.pct = 0.05,
  latent.vars = 'nCount_ATAC'
)
#1 DE peak 
Nalt_3_vs_meth_3 <- FindMarkers(
  object = results_cyto,
  ident.1 = 'Naltrexone_3',
  ident.2 = 'Methadone_3',
  only.pos = TRUE,
  test.use = 'LR',
  min.pct = 0.05,
  latent.vars = 'nCount_ATAC'
)


```


```{r session info}
devtools::session_info()
```