---
title: "Naive T 1 high resolution analysis"
author: "Jack Collora"
date: "2023-07-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis objectives 

Based on the previous results in the cumulative DE testing, it seems reasonable to focus on these cells. The specific quesitons to answer are:

1. What are the transcription factor activities that vary across conditions  
2. What are the proteins different across conditions  
3. What can we conclude about the transcriptional profiles different across individuals by  
 * GSEA  
 * GO analysis  
 
```{r loading in data libraries and GSEA functions}
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(fgsea))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(Seurat))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(msigdbr))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(Signac))
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(BiocParallel))
suppressPackageStartupMessages(library(enrichR))

register(MulticoreParam(4))

GSEA<-function(findmarks, genesets){
  findmarks$gene<-rownames(findmarks)
  genes<-findmarks %>%
    arrange(desc(avg_log2FC)) %>% 
    dplyr::select(gene, avg_log2FC)
  rnk<- deframe(genes)
  genes<- deframe(genes)
  genes<- fgsea(pathways=genesets, stats = genes,eps=FALSE)
  genes <- genes %>%
    as_tibble() %>%
    arrange(desc(NES))
  genes<-list(genes, rnk)
  names(genes)<-c("results","rnk")
  return(genes)
}

#Function augments the existing result table with pathway information, including the entire rank order list of gene hits and NES scores 
GSEATable<-function(GSEAwrap_out,gmt, gseaParam=1, name){
  #doing constants first since those are easy
  finalresults<-GSEAwrap_out[1][[1]]
  #first get the order of genes for each pathway, modified from the fGSEA plot function  
  rnk <- rank(GSEAwrap_out[2][[1]])
  ord <- order(rnk, decreasing = T)
  statsAdj <- GSEAwrap_out[2][[1]][ord]
  statsAdj <- sign(statsAdj) * (abs(statsAdj) ^ gseaParam)
  statsAdj <- statsAdj / max(abs(statsAdj))
  pathwaysetup<-list()
  NES<-list()
  print("start ranking")
  #basically here I just want to get everything into a character vector
  for (i in 1:length(finalresults$pathway)){
    pathway <- unname(as.vector(na.omit(match(gmt[[finalresults$pathway[[i]]]], names(statsAdj)))))
    pathway <- sort(pathway)
    NES[[i]]<-calcGseaStat(statsAdj, selectedStats = pathway,returnAllExtremes = TRUE)
    NES[[i]]<-NES[[i]]$tops
    pathwaysetup[[i]]<-pathway}
  print("done ranking")
  finalresults$rnkorder<-pathwaysetup
  finalresults$NESorder<-NES
  finalresults$ID<-name
  return(finalresults)
}


GSEAbig<-function(listofGSEAtables){
  GSEA<-bind_rows(listofGSEAtables)
  return(GSEA)
}



GSEAEnrichmentPlotComparison<-function(PathwayName, GSEACompOut, returnplot="none", cols=NA){
  top<-max(unlist(GSEACompOut$rnkorder))
  GSEACompOut<-GSEACompOut[GSEACompOut$pathway==PathwayName,]
  curves<-list()
  for (i in 1:length(rownames(GSEACompOut))){curves[[i]]<-list(c(0,GSEACompOut$rnkorder[[i]],top),c(0,GSEACompOut$NESorder[[i]],0))
  curves[[i]]<-as.data.frame(curves[[i]])
  curves[[i]]$name<-GSEACompOut$ID[[i]]
  names(curves[[i]])<-c("Rank","ES","Name")}
  curves<-bind_rows(curves)
  if(returnplot=="ES"){p<-EnrichmentScorePlot(curves)
  return(p)}
  if(returnplot=="BC"){p<-BarcodePlot(curves)
  return(p)}
  if(returnplot=="BOTH"){p<-ComboPlot(curves, Title = PathwayName, Table=GSEACompOut[GSEACompOut$pathway==PathwayName,c("ID","padj")], cols=cols)
  return(p)}
  return(curves)
}

BarcodePlot<-function(GSEACompPathway,stacked=FALSE){
  if(stacked==TRUE){GSEACompPathway<-split(GSEACompPathway,GSEACompPathway$Name)
  for (i in 1:length(GSEACompPathway)){GSEACompPathway[[i]]$y<-i}
  GSEACompPathway<-bind_rows(GSEACompPathway)
  GSEACompPathway$y<-((-GSEACompPathway$y)+1)
  p<-ggplot(GSEACompPathway,aes(x=Rank, y=y,xend=Rank, yend=y-1,color=Name))+geom_segment()+theme_classic()+theme(axis.title.y=element_blank(),
                                                                                                                  axis.text.y=element_blank(),
                                                                                                                  axis.ticks.y=element_blank())+xlab("Gene Rank") + xlim(-5,max(GSEACompPathway$Rank+100) )
  }else{
    p<-ggplot(GSEACompPathway,aes(x=Rank, y=-1,xend=Rank, yend=1,color=Name))+geom_segment()+theme_classic()
  }
  return(p)}


EnrichmentScorePlot<-function(GSEACompPathway){
  p<-ggplot(GSEACompPathway, aes(x=Rank, y=ES, color=Name))+geom_line(linewidth=1.5)+geom_hline(yintercept = 0)+theme_classic()+theme(axis.title.x=element_blank(),
                                                                                                                                 axis.text.x=element_blank(),
                                                                                                                                 axis.ticks.x=element_blank())+ylab("Enrichment Score")+xlim(-5,max(GSEACompPathway$Rank+100) )
  return(p)}
ComboPlot<-function(GSEACompPathway, Title="Enrichment Plot",Table=NA, cols=NA){
  EP<-EnrichmentScorePlot(GSEACompPathway)
  if(length(Table)>1){EP<-EP+annotation_custom(tableGrob(Table, theme = ttheme_minimal(), rows = NULL, cols = c("Name","q Value")), xmin = 10000, ymin = 0.60)}
  BC<-BarcodePlot(GSEACompPathway, stacked=TRUE)
  if(!is.na(cols[1])){
    BC<-BC+scale_color_manual(values = cols)
    EP<-EP+scale_color_manual(values = cols)
  }
  p<-ggarrange(EP,BC, common.legend = TRUE, ncol = 1, heights = c(0.6,0.25), align = "v", labels = Title)
  
  return(p)}


Meanenrichmentplot<-function(GSEAmain,greplist=c(),pathwaytodo=""){
    final<-list()
    IDssmooth<-paste(greplist, "smooth")
    IDsgray<-paste(greplist, "rough")
    #for each group, go through and make a smooth track 
    for(i in 1:length(greplist)){
    GSEAres<-GSEAmain[grepl(greplist[[i]], GSEAmain$ID) & grepl(pathwaytodo, GSEAmain$pathway),]
    GSEAres$split<-IDsgray[[i]]
    maximum<-max(unlist(GSEAres$rnkorder))
    #generate bins of size 5 (helps with smoothing)
    bins<-seq(1,maximum ,5)
    bins<-c(bins, maximum,maximum)
    res<-list()
    #average anypoint within those bins
    for (j in 1:(length(bins)-1)){
      res[[j]]<-mean(unlist(GSEAres$NESorder)[unlist(GSEAres$rnkorder)>bins[[j]]& unlist(GSEAres$rnkorder)<=bins[[j+1]]])
    }
    res<-c(res, 0)
    #remove any bin that does not have a point 
    bins<-bins[!is.na(res)]
    res<-res[!is.na(res)]
    #make a new track, add it to the overall 
    res<-list(GSEAres$pathway[[1]], NA, NA, NA, NA, NA, NA, NA, list(bins), list(res), "smooth", IDssmooth[[i]])
    GSEAres<-rbind(GSEAres, res)
    GSEAres$NESorder<-lapply(GSEAres$NESorder, as.double)
    curves<-list()
    #generate coordinate curves for each track including the smooth 
    for (j in 1:length(rownames(GSEAres))){curves[[j]]<-list(c(0,GSEAres$rnkorder[[j]],maximum),c(0,GSEAres$NESorder[[j]],0))
    curves[[j]]<-as.data.frame(curves[[j]])
    curves[[j]]$name<-GSEAres$ID[[j]]
    curves[[j]]$split<-GSEAres$split[[j]]
    names(curves[[j]])<-c("Rank","ES","Name","split")}
    final[[i]]<-bind_rows(curves)
    }
    #merge all the groups together
  final<-bind_rows(final)
  #plot things
  final<-ggplot(final[!grepl("smooth", final$split),],aes(x=Rank, y=ES, group=Name,color=split))+geom_line(linewidth=0.25)+geom_hline(yintercept = 0) + 
    geom_smooth(data = final[grepl("smooth", final$split),], aes(x=Rank, y=ES, group=split, color=split), linewidth=1.5, method = "gam")+theme_classic()+
    ggtitle(pathwaytodo)+ylab("Enrichment Score")+xlim(-5,max(maximum+100))
  
  return(final)}
 


GSEATable<-function(GSEAwrap_out,gmt, gseaParam=1, name){
  #doing constants first since those are easy
  finalresults<-GSEAwrap_out[1][[1]]
  #first get the order of genes for each pathway, modified from the fGSEA plot function  
  rnk <- rank(GSEAwrap_out[2][[1]])
  ord <- order(rnk, decreasing = T)
  statsAdj <- GSEAwrap_out[2][[1]][ord]
  statsAdj <- sign(statsAdj) * (abs(statsAdj) ^ gseaParam)
  statsAdj <- statsAdj / max(abs(statsAdj))
  pathwaysetup<-list()
  NES<-list()
  print("start ranking")
  #basically here I just want to get everything into a character vector
  for (i in 1:length(finalresults$pathway)){
    pathway <- unname(as.vector(na.omit(match(gmt[[finalresults$pathway[[i]]]], names(statsAdj)))))
    pathway <- sort(pathway)
    NES[[i]]<-calcGseaStat(statsAdj, selectedStats = pathway,returnAllExtremes = TRUE)
    NES[[i]]<-NES[[i]]$tops
    pathwaysetup[[i]]<-pathway}
  print("done ranking")
  finalresults$rnkorder<-pathwaysetup
  finalresults$NESorder<-NES
  finalresults$ID<-name
  return(finalresults)
}

FixMotifID<-function(markeroutput, seuratobj){
  markeroutput$genes<-ConvertMotifID(seuratobj, assay="ATAC", id=rownames(markeroutput))
  return(markeroutput)
}


GSEAbig<-function(listofGSEAtables){
  GSEA<-bind_rows(listofGSEAtables)
  return(GSEA)
}

VolPlot<-function(results,top=TRUE, adthresh=TRUE, thresh=0.2, threshn=30, Title=NA){
  results$value<-(-log10(results$p_val_adj))
  results$value[results$value==Inf]<-300
  results$gene<-rownames(results)
  p<-ggplot(results,aes(x=avg_log2FC, y=value, label=gene))+geom_point()+theme_classic()+xlab("Average Log Fold Change")+ylab("-log10 adjusted p value")+theme(plot.title = element_text(hjust = 0.5))
  if(adthresh==TRUE){thresh<-AdaptiveThreshold(results, Tstart=thresh,n=threshn)}
  if(top==TRUE){
    p<- p + geom_text_repel(aes(label=ifelse((avg_log2FC>thresh|avg_log2FC<(-thresh))&value>(-log10(0.01)) ,as.character(gene),'')),hjust=1,vjust=1)
  }
  if(!is.na(Title)){
    p<-p+ggtitle(Title)
  }
  return(p)
}
#quick selection of top n genes basically
AdaptiveThreshold<-function(results, n=30,Tstart=0.25){
  m=Inf
  while(n<=m){
    Tstart=Tstart+0.05
    m<-length(rownames(results)[(results$avg_log2FC>Tstart|results$avg_log2FC<(-Tstart))&results$value>10])
  }
  return(Tstart)
}

#borrowing some rushmore colors from the wes anderson color pack
BottleRocket2 = c("#FAD510", "#CB2314", "#273046", "#354823", "#1E1E1E")

#takes in enrichr output
PlotEnrichment<-function(data, topn=10, returnplot=TRUE, title="Significant Pathways"){
  data$value<- -log10(data$Adjusted.P.value)
  #order by most signfiicant 
  data$Term<-factor(data$Term, levels = rev(data$Term))
  data<-data[order(data$value, decreasing = TRUE),]
  #subset to top n 
  data<-data[1:topn,]
  plot<-ggplot(data, aes(x=value, y=Term, fill=ifelse(value>2, "Significant", "Not Significant")))+geom_bar(stat="identity")+xlab("-log10(p_adj)")+ylab("Pathway")+ggtitle(title)+
    theme_classic()+scale_fill_manual(values=c("Significant" = "#CB2314", "Not Significant" = "#1E1E1E"))+ guides(fill=guide_legend(title=""))
  if(returnplot){return(plot)}
  return(data)
}

#modded from WA pallette 
IsleofDogs1<-c("Bup.Nalo_vs_Nal_0" ="#9986A5", "Bup.Nalo_vs_Nal_3" = "#79402E", "Meth_vs_Bup.Nalo_0" = "#CCBA72", 
                 "Meth_vs_Bup.Nalo_3" = "#0F0D0E", "Meth_vs_Nal_0" = "#D9D0D3", "Meth_vs_Nal_3" = "#8D8680")


results<-readRDS("~/gibbs/DOGMAMORPH/Ranalysis/Objects/20230606FinalObj.rds") 

```
 
```{r genesets}

#grabbing hallmark as well as the curated, immune 
m_df_H<- msigdbr(species = "Homo sapiens", category = "H")
m_df_H<- rbind(msigdbr(species = "Homo sapiens", category = "C2"), m_df_H)
m_df_H<- rbind(msigdbr(species = "Homo sapiens", category = "C7"), m_df_H)
fgsea_sets<- m_df_H %>% split(x = .$gene_symbol, f = .$gs_name)
``` 

## transcription factor activities 

The objective is to see if there are specific transcription factors that may be driving differential activity across conditions. We will be comparing at time 0 and time 3 in a nonbias manner (DE) and the graphing specific factors of interest. 

Results here are mixed. Although JUN/FOS and related AP-1 signals are perturbed (similar to cytotoxic) they aren't as massively pertrubed as you might otherwise expect if this were exactly the same. 

```{r factor activities across conditions, fig.width=16, fig.height=9}
results$T_Tp<-paste(results$Treatment, results$Timepoint, sep = "_")
results$T_Tp<-factor(results$T_Tp, levels = c("Methadone_0","Methadone_3","Bup.Nalo_0", "Bup.Nalo_3", "Naltrexone_0", "Naltrexone_3"))
results_Naive_1<-subset(results, merged_clusters=="Naive_CD4_T_1")

DefaultAssay(results_Naive_1)<-"chromvar"
Idents(results_Naive_1)<-results_Naive_1$T_Tp

# month 3 comparisons 

Meth_vs_Nal_3<-FindMarkers(results_Naive_1, "Methadone_3","Naltrexone_3", mean.fxn = rowMeans)
Meth_vs_Bup.Nalo_3<-FindMarkers(results_Naive_1, "Methadone_3","Bup.Nalo_3", mean.fxn = rowMeans)
Bup.Nalo_vs_Nal_3<-FindMarkers(results_Naive_1, "Bup.Nalo_3","Naltrexone_3", mean.fxn = rowMeans)

# month 0 comparisons 
Meth_vs_Nal_0<-FindMarkers(results_Naive_1, "Methadone_0","Naltrexone_0", mean.fxn = rowMeans)
Meth_vs_Bup.Nalo_0<-FindMarkers(results_Naive_1, "Methadone_0","Bup.Nalo_0", mean.fxn = rowMeans)
Bup.Nalo_vs_Nal_0<-FindMarkers(results_Naive_1, "Bup.Nalo_0","Naltrexone_0", mean.fxn = rowMeans)


Meth_vs_Nal_3<-FixMotifID(Meth_vs_Nal_3, results_Naive_1)
Meth_vs_Bup.Nalo_3<-FixMotifID(Meth_vs_Bup.Nalo_3, results_Naive_1)
Bup.Nalo_vs_Nal_3<-FixMotifID(Bup.Nalo_vs_Nal_3, results_Naive_1)
Meth_vs_Nal_0<-FixMotifID(Meth_vs_Nal_0, results_Naive_1)
Meth_vs_Bup.Nalo_0<-FixMotifID(Meth_vs_Bup.Nalo_0, results_Naive_1)
Bup.Nalo_vs_Nal_0<-FixMotifID(Bup.Nalo_vs_Nal_0, results_Naive_1)

DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Bup.Nalo_vs_Nal_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_0)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_0)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Bup.Nalo_vs_Nal_0)


VlnPlot(results_Naive_1, "MA1573.2", pt.size = 0.1 )+ggtitle("Thap11")
VlnPlot(results_Naive_1, "MA0840.1", pt.size = 0.1 )+ggtitle("Creb5")
VlnPlot(results_Naive_1, "MA1133.1", pt.size = 0.1 )+ggtitle("JUN::JUNB")
VlnPlot(results_Naive_1, "MA0759.2", pt.size = 0.1 )+ggtitle("ELK3")
VlnPlot(results_Naive_1, "MA0764.3", pt.size = 0.1 )+ggtitle("ETV4")
VlnPlot(results_Naive_1, "MA1484.1", pt.size = 0.1 )+ggtitle("ETS2")
VlnPlot(results_Naive_1, "MA0834.1", pt.size = 0.1 )+ggtitle("ATF7")
VlnPlot(results_Naive_1, "MA1139.1", pt.size = 0.1 )+ggtitle("FOSL2::JUNB")
VlnPlot(results_Naive_1, "MA1129.1", pt.size = 0.1 )+ggtitle("FOSL1::JUN")
VlnPlot(results_Naive_1, "MA0056.2", pt.size = 0.1 )+ggtitle("MZF1")
VlnPlot(results_Naive_1, "MA0101.1", pt.size = 0.1 )+ggtitle("RELA")
VlnPlot(results_Naive_1, "MA0081.2", pt.size = 0.1 )+ggtitle("SPIB")
VlnPlot(results_Naive_1, "MA0156.3", pt.size = 0.1 )+ggtitle("FEV")
VlnPlot(results_Naive_1, "MA0139.1", pt.size = 0.1 )+ggtitle("CTCF")
VlnPlot(results_Naive_1, "MA0645.1", pt.size = 0.1 )+ggtitle("ETV6")
VlnPlot(results_Naive_1, "MA0081.2", pt.size = 0.1 )+ggtitle("SPIB")
VlnPlot(results_Naive_1, "MA1512.1", pt.size = 0.1 )+ggtitle("KLF11")


Idents(results_Naive_1)<-factor(Idents(results_Naive_1),levels = c("Methadone_0","Bup.Nalo_0","Naltrexone_0","Methadone_3", "Bup.Nalo_3", "Naltrexone_3"))

VlnPlot(results_Naive_1, "MA1573.2", pt.size = 0.1 )+ggtitle("Thap11")
VlnPlot(results_Naive_1, "MA0840.1", pt.size = 0.1 )+ggtitle("Creb5")
VlnPlot(results_Naive_1, "MA1133.1", pt.size = 0.1 )+ggtitle("JUN::JUNB")
VlnPlot(results_Naive_1, "MA0759.2", pt.size = 0.1 )+ggtitle("ELK3")
VlnPlot(results_Naive_1, "MA0764.3", pt.size = 0.1 )+ggtitle("ETV4")
VlnPlot(results_Naive_1, "MA1484.1", pt.size = 0.1 )+ggtitle("ETS2")
VlnPlot(results_Naive_1, "MA0834.1", pt.size = 0.1 )+ggtitle("ATF7")
VlnPlot(results_Naive_1, "MA1139.1", pt.size = 0.1 )+ggtitle("FOSL2::JUNB")
VlnPlot(results_Naive_1, "MA1129.1", pt.size = 0.1 )+ggtitle("FOSL1::JUN")
VlnPlot(results_Naive_1, "MA0056.2", pt.size = 0.1 )+ggtitle("MZF1")
VlnPlot(results_Naive_1, "MA0101.1", pt.size = 0.1 )+ggtitle("RELA")
VlnPlot(results_Naive_1, "MA0081.2", pt.size = 0.1 )+ggtitle("SPIB")
VlnPlot(results_Naive_1, "MA0156.3", pt.size = 0.1 )+ggtitle("FEV")
VlnPlot(results_Naive_1, "MA0139.1", pt.size = 0.1 )+ggtitle("CTCF")
VlnPlot(results_Naive_1, "MA0645.1", pt.size = 0.1 )+ggtitle("ETV6")
VlnPlot(results_Naive_1, "MA0081.2", pt.size = 0.1 )+ggtitle("SPIB")
VlnPlot(results_Naive_1, "MA1512.1", pt.size = 0.1 )+ggtitle("KLF11")

```

## Surface protein differences 

Doing the same thing as Chormvar but instead we're looking at protein expression. 

As below, this cell type basically has no differences in protein expression across conditions. No significant factor is identified. 

```{r proteins across conditions, fig.width=16, fig.height=9}

DefaultAssay(results_Naive_1)<-"ADT"
Idents(results_Naive_1)<-results_Naive_1$T_Tp

#need to first do CLR by library 

results_Naive_1_split<-SplitObject(results_Naive_1, split.by = "orig.ident")

results_Naive_1_split<-lapply(results_Naive_1_split, NormalizeData, normalization.method = "CLR")

for (i in 1:9){
  results_Naive_1_split[[i]]<-results_Naive_1_split[[i]]@assays$ADT@data
}
results_Naive_1_split<-do.call(cbind, results_Naive_1_split)
#put it in but ensure things are in the correct order
results_Naive_1@assays$ADT@data<-results_Naive_1_split[,colnames(results_Naive_1)]

Meth_vs_Nal_3<-FindMarkers(results_Naive_1, "Methadone_3","Naltrexone_3")
Meth_vs_Bup.Nalo_3<-FindMarkers(results_Naive_1, "Methadone_3","Bup.Nalo_3")
Bup.Nalo_vs_Nal_3<-FindMarkers(results_Naive_1, "Bup.Nalo_3","Naltrexone_3")

# month 0 comparisons 
Meth_vs_Nal_0<-FindMarkers(results_Naive_1, "Methadone_0","Naltrexone_0")
Meth_vs_Bup.Nalo_0<-FindMarkers(results_Naive_1, "Methadone_0","Bup.Nalo_0")
Bup.Nalo_vs_Nal_0<-FindMarkers(results_Naive_1, "Bup.Nalo_0","Naltrexone_0")

#Bup.Nalo and Nal are not different at all (0 pass logfc)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_0)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_0)



```

## Gene expression and GSEA 


### Standard DE 

We're just using the standard wilcox/mann whitney U test here to look at the impact of these different conditions on gene expression
```{r standard DE, fig.width=16, fig.height=9}
DefaultAssay(results_Naive_1)<-"RNA"

# month 3 comparisons 

Meth_vs_Nal_3<-FindMarkers(results_Naive_1, "Methadone_3","Naltrexone_3")
Meth_vs_Bup.Nalo_3<-FindMarkers(results_Naive_1, "Methadone_3","Bup.Nalo_3")
Bup.Nalo_vs_Nal_3<-FindMarkers(results_Naive_1, "Bup.Nalo_3","Naltrexone_3")

# month 0 comparisons 
Meth_vs_Nal_0<-FindMarkers(results_Naive_1, "Methadone_0","Naltrexone_0")
Meth_vs_Bup.Nalo_0<-FindMarkers(results_Naive_1, "Methadone_0","Bup.Nalo_0")
Bup.Nalo_vs_Nal_0<-FindMarkers(results_Naive_1, "Bup.Nalo_0","Naltrexone_0")


DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Bup.Nalo_vs_Nal_3)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Nal_0)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Meth_vs_Bup.Nalo_0)
DT::datatable(rownames=TRUE, filter="top", class='cell-border stripe', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), data=Bup.Nalo_vs_Nal_0)

VlnPlot(results_Naive_1, "KLRG1", pt.size = 0.1)

#volcanos of just those highly differentially expressed
VolPlot(Meth_vs_Nal_3)
VolPlot(Meth_vs_Bup.Nalo_3)
VolPlot(Bup.Nalo_vs_Nal_3)
VolPlot(Meth_vs_Nal_0)
VolPlot(Meth_vs_Bup.Nalo_0)
VolPlot(Bup.Nalo_vs_Nal_0)

#doing it again but with no cutoffs for FC

# month 3 comparisons 

Meth_vs_Nal_3<-FindMarkers(results_Naive_1, "Methadone_3","Naltrexone_3", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)
Meth_vs_Bup.Nalo_3<-FindMarkers(results_Naive_1, "Methadone_3","Bup.Nalo_3", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)
Bup.Nalo_vs_Nal_3<-FindMarkers(results_Naive_1, "Bup.Nalo_3","Naltrexone_3", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)

# month 0 comparisons 
Meth_vs_Nal_0<-FindMarkers(results_Naive_1, "Methadone_0","Naltrexone_0", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)
Meth_vs_Bup.Nalo_0<-FindMarkers(results_Naive_1, "Methadone_0","Bup.Nalo_0", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)
Bup.Nalo_vs_Nal_0<-FindMarkers(results_Naive_1, "Bup.Nalo_0","Naltrexone_0", min.pct = -Inf, min.diff.pct = -Inf, logfc.threshold = -Inf)

#not placing data tables for these very large dataframes as they make the resulting html unusable 

VolPlot(Meth_vs_Nal_3)
VolPlot(Meth_vs_Bup.Nalo_3)
VolPlot(Bup.Nalo_vs_Nal_3)
VolPlot(Meth_vs_Nal_0)
VolPlot(Meth_vs_Bup.Nalo_0)
VolPlot(Bup.Nalo_vs_Nal_0)
```
#### GO analysis of DE genes from traditional analyses 

Using Enrichr we can look for genesets that are enriched here for each comparison. 

In general, the pathways here are interesting because they suggest a disregulation of certain activation pathways sepcific to some conditions over others. 

```{r GO_trad, fig.width=16, fig.height=9}
#outputing the genelists to test 



to_output<-list(Meth_vs_Nal_3,Meth_vs_Bup.Nalo_3,Bup.Nalo_vs_Nal_3,Meth_vs_Nal_0,Meth_vs_Bup.Nalo_0,Bup.Nalo_vs_Nal_0 )
names(to_output)<-c("Meth_vs_Nal_3","Meth_vs_Bup.Nalo_3","Bup.Nalo_vs_Nal_3","Meth_vs_Nal_0","Meth_vs_Bup.Nalo_0","Bup.Nalo_vs_Nal_0")
pcut<-0.05 
LFCcut<-0.25

websiteLive <- getOption("enrichR.live")
dbs<-c("GO_Biological_Process_2023","GO_Molecular_Function_2023")
outdir<-"~/gibbs/DOGMAMORPH/Ranalysis/enrichr/Trad_DE_Naive_1"
#basically we'll loop through, write the subset of genes based on those cutoffs 
for (i in 1:length(to_output)){
  cur<-to_output[[i]][(to_output[[i]]$avg_log2FC< -LFCcut) | (to_output[[i]]$avg_log2FC> LFCcut),]
  cur$p_val_adj<-p.adjust(cur$p_val, "BH")
  up<-rownames(cur)[(cur$p_val_adj<pcut)&(cur$avg_log2FC>0)]
  down<-rownames(cur)[(cur$p_val_adj<pcut)&(cur$avg_log2FC<0)]
  write.table(up, paste0(outdir,"/", names(to_output)[[i]], "_up.txt"), 
              quote = FALSE, row.names = FALSE, col.names = FALSE)
  write.table(down, paste0(outdir,"/", names(to_output)[[i]], "_down.txt"), 
              quote = FALSE, row.names = FALSE, col.names = FALSE)
  up<-enrichr(up, dbs)
  up_proc<-cbind(up$GO_Biological_Process_2023)
  up_func<-cbind(up$GO_Molecular_Function_2023)

  down<-enrichr(down, dbs)
  down_proc<-cbind(down$GO_Biological_Process_2023)
  down_func<-cbind(down$GO_Molecular_Function_2023)

  print(PlotEnrichment(up_proc, topn = 20, title=paste("GO_proc",names(to_output)[[i]], "up" ,sep="_")))
  print(PlotEnrichment(down_proc, topn = 20, title=paste("GO_proc",names(to_output)[[i]], "down" ,sep="_")))
  print(PlotEnrichment(up_func, topn = 20, title=paste("GO_func",names(to_output)[[i]], "up" ,sep="_")))
  print(PlotEnrichment(down_func, topn = 20, title=paste("GO_func",names(to_output)[[i]], "down" ,sep="_")))
  
  write.table(up_proc,
              paste0("~/gibbs/DOGMAMORPH/Ranalysis/enrichr/Trad_DE_Naive_1/Results_tables/","GO_proc_",names(to_output)[[i]], "_up.txt" ,sep="_"), quote = FALSE, row.names = FALSE)
  write.table(down_proc,
              paste0("~/gibbs/DOGMAMORPH/Ranalysis/enrichr/Trad_DE_Naive_1/Results_tables/","GO_proc_",names(to_output)[[i]], "_down.txt" ,sep="_"), quote = FALSE, row.names = FALSE)
  write.table(up_func,
              paste0("~/gibbs/DOGMAMORPH/Ranalysis/enrichr/Trad_DE_Naive_1/Results_tables/","GO_func_",names(to_output)[[i]], "_up.txt" ,sep="_"), quote = FALSE, row.names = FALSE)
  write.table(down_func, paste0("~/gibbs/DOGMAMORPH/Ranalysis/enrichr/Trad_DE_Naive_1/Results_tables/","GO_func_",names(to_output)[[i]], "_down.txt" ,sep="_"), sep = "\t", quote = FALSE, row.names = FALSE)
  }




```
#### GSEA analysis of DE genes from traditional analyses 

GSEA is a nice addition particularly in single cell because the signals are weaker due to inherent sparsity and pooling info across genes by ranking can be more powerful. 

```{r GSEA_trad, fig.width=16, fig.height=9}
GSEAres<-list()
for (i in 1:length(to_output)){
GSEAres[[i]]<-GSEA(to_output[[i]], genesets = fgsea_sets)
GSEAres[[i]]<-GSEATable(GSEAwrap_out =GSEAres[[i]], gmt = fgsea_sets, name = names(to_output)[[i]] )}
GSEAres<-GSEAbig(listofGSEAtables = GSEAres)

saveRDS(GSEAres, file="~/gibbs/DOGMAMORPH/Ranalysis/Objects/20230711GSEAnaive1res.rds")
#identified these in 20230627cytotoxicGSEA.R

to_test<-c("REACTOME_EUKARYOTIC_TRANSLATION_ELONGATION","REACTOME_G_ALPHA_12_13_SIGNALLING_EVENTS","GSE10325_LUPUS_CD4_TCELL_VS_LUPUS_BCELL_DN",
           "GSE10325_LUPUS_CD4_TCELL_VS_LUPUS_MYELOID_DN","HALLMARK_INTERFERON_GAMMA_RESPONSE","GSE26495_NAIVE_VS_PD1HIGH_CD8_TCELL_DN",
           "GSE26495_NAIVE_VS_PD1LOW_CD8_TCELL_DN","GSE41978_ID2_KO_VS_ID2_KO_AND_BIM_KO_KLRG1_LOW_EFFECTOR_CD8_TCELL_DN","GSE10239_NAIVE_VS_KLRG1HIGH_EFF_CD8_TCELL_UP",
           "GSE7460_CD8_TCELL_VS_CD4_TCELL_ACT_UP","GSE9650_NAIVE_VS_EFF_CD8_TCELL_DN","GSE3039_NKT_CELL_VS_ALPHAALPHA_CD8_TCELL_DN",
           "GSE33424_CD161_INT_VS_NEG_CD8_TCELL_UP","GSE21670_UNTREATED_VS_IL6_TREATED_STAT3_KO_CD4_TCELL_DN","GSE21670_UNTREATED_VS_TGFB_IL6_TREATED_CD4_TCELL_DN",
           "GSE2770_TGFB_AND_IL4_ACT_VS_ACT_CD4_TCELL_2H_UP","GSE39820_CTRL_VS_IL1B_IL6_IL23A_CD4_TCELL_UP","GSE2770_IL12_AND_TGFB_ACT_VS_ACT_CD4_TCELL_48H_DN",
           "REACTOME_INFLUENZA_INFECTION","GSE34205_HEALTHY_VS_FLU_INF_INFANT_PBMC_UP","HOWARD_MONOCYTE_INACT_MONOV_INFLUENZA_A_INDONESIA_05_2005_H5N1_AGE_18_49YO_1DY_UP",
           "REACTOME_HIV_INFECTION","HALLMARK_INTERFERON_ALPHA_RESPONSE","ZHANG_INTERFERON_RESPONSE",
           "GSE24574_BCL6_HIGH_TFH_VS_TFH_CD4_TCELL_DN","GSE26030_TH1_VS_TH17_RESTIMULATED_DAY5_POST_POLARIZATION_UP","GSE3982_MEMORY_CD4_TCELL_VS_TH1_DN",
           "PHONG_TNF_RESPONSE_NOT_VIA_P38","WANG_TNF_TARGETS","ZHOU_TNF_SIGNALING_4HR","PHONG_TNF_TARGETS_UP","GSE11057_CD4_EFF_MEM_VS_PBMC_UP")

for (i in 1:length(to_test)){
  print(GSEAEnrichmentPlotComparison(to_test[i], GSEAres, "BOTH", cols = IsleofDogs1))
}

```



```{r session info}
devtools::session_info()
```